### 2024-04-25 data/data-prep2.qmd

---
title: "data-prep2"
format: html
---
```{r}
rm(list = ls())
```

Let's work with the TTM that we already have ! Only destinations within 15 minutes or less, for walking mode.
```{r library-setup}
library(biscale)
library(cowplot)
library(sf)
library(here)
library(dplyr)
library(tmap)
library(mapview)
library(tmaptools)
library(pander)
library(flextable)
library(ftExtra)
library(scales)
```

```{r spatial-availability-function}
sp_avail_detailed <- function(x, o_id, d_id, pop, opp, r, f, alpha = 1){

  o_id <- rlang::enquo(o_id)
  d_id <- rlang::enquo(d_id)
  pop <- rlang::enquo(pop)
  opp <- rlang::enquo(opp)
  r <- rlang::enquo(r)
  f <- rlang::enquo(f)

  sum_pop <- x |>
    dplyr::distinct(!!o_id,
                    .keep_all = TRUE) |>
    dplyr::mutate(sum_pop = !!r*(!!pop)^alpha) |>
    dplyr::pull(sum_pop) |>
    sum()

  f_p <- dplyr::pull(x, !!r) * dplyr::pull(x, !!pop)^alpha / sum_pop

  sum_impedance <- x |>
    dplyr::group_by(!!d_id) |>
    dplyr::summarize(sum_impedance = sum(!!f))

  x <- x |>
    dplyr::left_join(sum_impedance,
                     by = rlang::as_name(d_id))

  f_c <- dplyr::pull(x, !!f) / x$sum_impedance

  x$f_c <- f_c
  x$f_p <- f_p

  sum_pa <- x |>
    dplyr::group_by(!!d_id) |>
    dplyr::summarize(sum_pa= sum(f_p * f_c))

  x <- x |>
    dplyr::left_join(sum_pa,
                     by = rlang::as_name(d_id))
  x$f_t <- (f_p * f_c) / dplyr::pull(x, sum_pa)

  x |>
    dplyr::mutate(V_ij = !!opp * f_t)
}
```

```{r}
#boundaries for mapping
Community_Boundaries <- st_read("data-raw/Community_Boundaries.shp") |> mutate(COMMUNITY_ = ifelse(COMMUNITY_ == "Hamilton", "Hamilton-Central", COMMUNITY_))|> st_transform(crs=4326)
City_Boundary <- st_read("data-raw/City_Boundary.shp") |> st_transform(crs=4326)
hydro_p_LakeOntario <- st_read("data-raw/hydro_p_LakeOntario.shp") |> st_transform(crs=4326)
ham_bay <- st_read("data-raw/Waterbodies.shp") |> st_transform(crs=4326) |> filter(FEATURE_TY == "Lake")
ham_bay_cropped<-st_crop(ham_bay$geometry, Community_Boundaries$geometry)
hydro_p_LakeOntario_cropped<-st_crop(hydro_p_LakeOntario$geometry, Community_Boundaries$geometry)

load("care_dest.rda")
load("ttm_care_walk.rda")
load("HAM_census_21.rda")
```

```{r}
HAM_census_21 <- HAM_census_21 |>
  rename("LICO_AT_TOTAL" = `v_CA21_1085: Prevalence of low income based on the Low-income cut-offs, after tax (LICO-AT) (%)`) |> 
  mutate(LICO_AT_TOTAL = ifelse(is.na(LICO_AT_TOTAL), 0, LICO_AT_TOTAL)) |>
  rowwise() |> mutate(
         perc_walked = (`v_CA21_7647: Walked`)/sum(`v_CA21_7647: Walked`,`v_CA21_7635: Car, truck or van`, `v_CA21_7644: Public transit`,`v_CA21_7650: Bicycle`, `v_CA21_7653: Other method`, na.rm=TRUE),
         perc_car = (`v_CA21_7635: Car, truck or van`)/sum(`v_CA21_7647: Walked`,`v_CA21_7635: Car, truck or van`, `v_CA21_7644: Public transit`,`v_CA21_7650: Bicycle`, `v_CA21_7653: Other method`, na.rm=TRUE),
perc_transit = (`v_CA21_7644: Public transit`)/sum(`v_CA21_7647: Walked`,`v_CA21_7635: Car, truck or van`, `v_CA21_7644: Public transit`,`v_CA21_7650: Bicycle`, `v_CA21_7653: Other method`, na.rm=TRUE),
 perc_cycle_other = (`v_CA21_7650: Bicycle`+ `v_CA21_7653: Other method`)/sum(`v_CA21_7647: Walked`,`v_CA21_7635: Car, truck or van`, `v_CA21_7644: Public transit`,`v_CA21_7650: Bicycle`, `v_CA21_7653: Other method`, na.rm=TRUE),
pop_den = Population/`Shape Area`) |>
  select(c("GeoUID", "Population", "LICO_AT_TOTAL", "perc_walked", "perc_car", "perc_transit", "perc_cycle_other", "pop_den"))

HAM_census_21 <- st_join(HAM_census_21, Community_Boundaries|> select("COMMUNITY_"), largest=TRUE)
DA_centroids <- HAM_census_21 |> st_centroid() 
```

```{r reformat care-dest specific categories}
care_dest <- care_dest |> 
  mutate(Care_Category_Specific = case_when(
    Care_Category_Specific  == "Daycare_EarlyON" ~ "Daycare or EarlyON",
    Care_Category_Specific == "Community_Centre" ~ "Community or Recreation Centre",
    Care_Category_Specific == "Daycare_Centre" ~ "Daycare or EarlyON",
    Care_Category_Specific == "Recreation_Centre" ~ "Community or Recreation Centre",
    Care_Category_Specific == "School" ~ "School",
    Care_Category_Specific == "Long-Term_Care_Home" ~ "Long-Term Care or Retirement Home",
    Care_Category_Specific == "Retirement_Home" ~ "Long-Term Care or Retirement Home",
    Care_Category_Specific == "Senior_Centre" ~ "Senior Centre",
    Care_Category_Specific == "Seniors_Active_Living_Centre" ~ "Senior Centre",
    Care_Category_Specific == "Post_Office" ~ "Post Office",
    Care_Category_Specific == "Library" ~ "Library",
    Care_Category_Specific == "Banks" ~ "Bank or ATM",
    Care_Category_Specific == "Grocer_ConvenienceStore" ~ "Convenience Store",
    Care_Category_Specific == "Clinics" ~ "Hospital or Clinic",
    Care_Category_Specific == "Grocer_Retail" ~ "Grocer",
    Care_Category_Specific == "Dentist" ~ "Dentist",
    Care_Category_Specific == "Hosptial" ~ "Hospital or Clinic",
    Care_Category_Specific == "Hospital" ~ "Hospital or Clinic",
    Care_Category_Specific == "Pharmacy" ~ "Pharmacy",
    Care_Category_Specific == "Park" ~ "Park"),
  Care_Category = case_when(
    Care_Category == "Child_Centric" ~ "Dependent-centric",
    Care_Category == "Elder_Centric" ~ "Dependent-centric",
    Care_Category == "Errand_Centric" ~ "Errand-centric",
    Care_Category == "Grocery_Centric" ~ "Grocery-centric",
    Care_Category == "Health_Centric" ~ "Health-centric"))
```

```{r travel-time-15min}
#15 mins or less
ttm_care_walk_15 <- ttm_care_walk |> 
  filter(travel_time_p50 <= 15) |> select(-c("travel_time_p25", "travel_time_p75")) |> 
  left_join(
  care_dest |> st_drop_geometry() |> select(c("ID", "Care_Category", "Care_Category_Specific", "Care_Category_Speciality")) |>
    mutate(ID = as.character(ID)),
  by = c("to_id" = "ID")) |>
  mutate(f = 1)

ttm_care_walk_15 <- ttm_care_walk_15 |>
  left_join(HAM_census_21 |> st_drop_geometry() |> select(c("GeoUID", "Population", "LICO_AT_TOTAL")), by=c("from_id" = "GeoUID")) 
```

```{r}
access_15 <- ttm_care_walk_15 |> group_by(from_id,Care_Category, Care_Category_Specific) |> summarize(copp_isc = n())

test_all <- ttm_care_walk_15 |>
  sp_avail_detailed(o_id = from_id,
                    d_id = to_id, 
                    pop = Population, 
                    opp = f,
                    r = f, 
                    f = f, alpha = 1) |> group_by(from_id, Care_Category, Care_Category_Specific) |> summarize(V_isc = sum(V_ij, na.rm = TRUE))

access_15 <- access_15 |> left_join(test_all, by=c("from_id", "Care_Category", "Care_Category_Specific")) |> ungroup()
rm(test_all)
access_15$V_isc |> sum(na.rm = TRUE) #sums to 1997 opportunities -- out of 2225 total opportunities
```

```{r}
# all speciality categories should be represented, i.e., each ID should have 14 rows -- if that category is empty it should be 0. Let's add those 0s. 
access_15_zeros <- 
  data.frame(from_id = rep(HAM_census_21$GeoUID, length(access_15$Care_Category_Specific |> unique())),
           Care_Category = rep(c("Dependent-centric", "Dependent-centric", "Grocery-centric", "Dependent-centric", 
                                 "Dependent-centric", "Dependent-centric",
                                 "Errand-centric", "Errand-centric", 
                                 "Grocery-centric","Health-centric",
                                 "Health-centric", "Errand-centric",
                                 "Dependent-centric", "Health-centric"), length(HAM_census_21$GeoUID)),

           Care_Category_Specific = rep(c("Park", "Daycare or EarlyON", "Convenience Store", "School",
                                          "Long-Term Care or Retirement Home", "Senior Centre",
                                          "Bank or ATM", "Library", 
                                          "Grocer", "Dentist", 
                                          "Pharmacy", "Post Office",
                                          "Community or Recreation Centre", "Hospital or Clinic"), length(HAM_census_21$GeoUID)),

           copp_isc = rep(0, length(HAM_census_21$GeoUID)*length(c("Park", "Daycare or EarlyON", "Convenience Store", "School",
                                          "Long-Term Care or Retirement Home", "Senior Centre",
                                          "Bank or ATM", "Library", 
                                          "Grocer", "Dentist", 
                                          "Pharmacy", "Post Office",
                                          "Community or Recreation Centre", "Hospital or Clinic"))),

           V_isc = rep(0, length(HAM_census_21$GeoUID)*length(c("Park", "Daycare or EarlyON", "Convenience Store", "School",
                                          "Long-Term Care or Retirement Home", "Senior Centre",
                                          "Bank or ATM", "Library", 
                                          "Grocer", "Dentist", 
                                          "Pharmacy", "Post Office",
                                          "Community or Recreation Centre", "Hospital or Clinic"))))

access_15 <- bind_rows(access_15_zeros, access_15) |>
                         group_by(from_id, Care_Category, Care_Category_Specific) |> 
  summarize(copp_isc = sum(copp_isc),
            V_isc = sum(V_isc))

access_15_bycategory <- access_15 |> group_by(from_id, Care_Category)  |> 
  summarize(copp_ic = sum(copp_isc, na.rm=T),
            V_ic = sum(V_isc, na.rm = T))

access_15 <- access_15 |>
  left_join(HAM_census_21 |> st_drop_geometry() |> select(c("GeoUID", "Population", "LICO_AT_TOTAL", "COMMUNITY_")), 
            by=c("from_id" = "GeoUID")) |> 
  mutate(copp_isc_perpop = copp_isc/Population,
         v_isc = V_isc/Population) |> ungroup()

access_15_bycategory <- access_15_bycategory |>
  left_join(HAM_census_21 |> st_drop_geometry() |> select(c("GeoUID", "Population", "LICO_AT_TOTAL", "COMMUNITY_")), 
            by=c("from_id" = "GeoUID")) |> 
  mutate(copp_ic_perpop = copp_ic/Population,
         v_ic = V_ic/Population) |> ungroup()

rm(access_15_zeros)
```
```{r}
# access_15
access_15 |> select(c(Care_Category, Care_Category_Specific)) |> unique()
# access_15 |> filter(Care_Category == c("Errand-centric", "Grocery-centric","Health-centric"))
```

```{r}
no_dependents <- access_15_bycategory |> filter(Care_Category == "Errand-centric"|
                                                 Care_Category == "Grocery-centric" |
                                                  Care_Category == "Health-centric") |> 
  group_by(from_id) |>
  summarise(Care_Access_Status = case_when(all(copp_ic == 0) ~ "Not caring",
                                 all(copp_ic >= 1 & copp_ic <= 2) ~ "Caring", 
                                 all(copp_ic >= 3) ~ "Very caring"),
            copp_isc = sum(copp_ic), 
            V_isc = sum(V_ic),
            Population = first(Population),
            COMMUNITY_ = first(COMMUNITY_))

no_dependents2 <- access_15 |> filter(Care_Category == "Errand-centric"|
                                                 Care_Category == "Grocery-centric" |
                                                  Care_Category == "Health-centric") |> 
  group_by(from_id) |>
  summarise(Care_Access_Status2 = case_when(all(copp_isc >= 1 & copp_isc < 2) ~ "Completely caring", 
                                 all(copp_isc >= 2) ~ "Very completely caring"))

no_dependents <- no_dependents |> left_join(no_dependents2, by= c("from_id"))

```

```{r}
no_dependents
```

<!-- NOTES FROM LEA CONVO FEB 22:

We changed some category tags. 
 - child and elder centric are now both 'dependent-centric'. These are 'caring for others'. We also consolidated some sub-categories. Daycare and EarlyON are now combined.. Long Term-Care Home and Residential Home are now combined.. Senior Centre and Senior Active Living Centre are combined. Hospital and Clinic are combined. Recreation Centre and COmmunity Centre are now combined.
 - A consideration for the future * maybe convenience stores and grocery stores can be be modulated based on their time -- i.e., if a DA trip to convenience is less than grocery, it is counted, if it is more than grocery it is not. More thought needs to be put into this, for now I did not include this consideration. 
 - We will consider the following 2 (or maybe 3?) faceted (by mode) map with scales:

1.	No-children (oneself and partner, no dependents) city map (15 min walk + 15 min cycle):
-	Errand, grocery, health

-	Scale: 
o	not-caring (does not have 1 of each category), 
o	caring (1 of each category), 
o	very caring (2 or more of each category), 
o	completely caring (1 of each sub-category), 
o	very completely caring (2 of each sub-category)

2.	Children+seniors: -- the truly caring city has all 4 categories. You always preform care but not always care for others. Your status will change. 
-	+elder+children
-	Scale: (same)

3. What neighbourhoods are ‘almost there’ for being completely caring? What are they missing?

And our presentation's angle: a 15 minute city is XYZ, the destinations are contested but typically are 'care' trips as well as discretionary trips (fun stuff). The mobility of care framework: paid, unpaid, and discretionary. All is important. But unpaid travel is under-studied. We are going to explore unpaid trips: they are essential and short-distance trips. A masculine city is one where paid work + discretionary trips are planned for. This analysis presents a tool to conceptual what is and is not a 'caring' or 'caring complete' neighbourhood. presents some visualisations.

APRIL 25 2024
- for simplicity, 15 min walk thresholds as a baseline for 15-min caring neighbourhoods. 15 min cycle thresholds to demonstrate the potential of cyling to get those 'almost there' neighbourhoods to be more caring if cycling infrastructure is implemented. Transit/bike to care isn't so common currently, but it's ane equity argument and shows political potential.

-->
<!-- THINK FROM ABOVE HERE-->

```{r message=FALSE}
all_cats <- access_15 |> group_by(from_id) |>
  summarise(Complete = ifelse(all(c_opp > 0 ), 1, 0),
            c_opp = sum(c_opp), 
            V_ic = sum(V_ic),
            Population = first(Population),
            COMMUNITY_ = first(COMMUNITY_))

all_cats |> group_by(COMMUNITY_) |> 
  summarize(DApop_served_all_categories = sum(Population*Complete, na.rm=T),
            DApopperc_served_all_categories = sum(Population*Complete, na.rm=T)/sum(Population))

one_cat <- access_15 |> group_by(from_id, Care_Category)|>
  summarise(Complete = ifelse(all(c_opp > 0 ), 1, 0),
            c_opp = sum(c_opp), 
            V_ic = sum(V_ic),
            Population = first(Population),
            COMMUNITY_ = first(COMMUNITY_))

one_cat_communities <- one_cat |> group_by(COMMUNITY_, Care_Category) |> 
  summarize(DApop_served_1_category = sum(Population*Complete, na.rm=T),
            DApopperc_served_1_category = sum(Population*Complete, na.rm=T)/sum(Population))

one_cat_communities |> filter(COMMUNITY_ == "Hamilton-Central")
one_cat_communities |> filter(Care_Category == "Grocery_Centric") 
```
```{r}
all_cats <- left_join(HAM_census_21 |> select(c(GeoUID)), all_cats, by=c( "GeoUID" = "from_id"))
```
```{r message=FALSE}
tmap_mode("view")
tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(all_cats) +
  tm_polygons("c_opp",
              palette = "Greens",
              border.alpha = 0.2,
              title = "Number of destinations spatially accessible within 15 mins.",
              legend.is.portrait = FALSE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 20)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12,
            legend.title.size = 0.7,
            legend.text.size = 0.6)
```

```{r message=FALSE}
tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(all_cats) +
  tm_polygons("V_ic",
              palette = "Purples",
              border.alpha = 0,
              title = "Spatially available destinations to mode-using population within 15 mins.",
              legend.is.portrait = FALSE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 20)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12,
            legend.title.size = 0.7,
            legend.text.size = 0.6)

```
```{r}
one_cat <- left_join(HAM_census_21 |> select(c(GeoUID)), one_cat, by=c( "GeoUID" = "from_id"))
```

```{r}
tmap_mode("plot")
tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(one_cat) +
  tm_polygons("Complete",
              border.alpha = 0,
              legend.is.portrait = FALSE,
              n=2,
              labels = c("Not complete", "Complete")) +
  tm_facets(by=c("Care_Category"), nrow=2, showNA = FALSE)+
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 20)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12,
            legend.title.size = 0.7,
            legend.text.size = 0.6)
```

Questions:

- what is a reasonable travel time? Is 15 minute walk indeed fine for our purpose?

- is 1 potential opportunity enough per subcategory?

- Should some subcategories be substitutes? E.g., 1 daycare OR EARLYON centre is enough, not 1 of each.

- Are there ways the categories can overlap to tell a story? Maybe to include google maps images. eg. 
  - communities where children+errand+grocery centric care is easier -- especially for women
  - health_centric+Elder_centric -- people with higher health care needs

- care trips 30% of all trips made. they are essential and have

<!-- ```{r} -->
<!-- test1 <- ttm_care_walk_15 |> filter(Care_Category == "Child_Centric") |> -->
<!--   sp_avail_detailed(o_id = from_id, -->
<!--                     d_id = to_id,  -->
<!--                     pop = Population,  -->
<!--                     opp = f, -->
<!--                     r = f,  -->
<!--                     f = f, alpha = 1)|> -->
<!--   group_by(from_id, Care_Category) |> summarize(V_ic = sum(V_ij, na.rm = TRUE))  -->

<!-- test2 <- ttm_care_walk_15 |> filter(Care_Category == "Grocery_Centric") |> -->
<!--   sp_avail_detailed(o_id = from_id, -->
<!--                     d_id = to_id,  -->
<!--                     pop = Population,  -->
<!--                     opp = f, -->
<!--                     r = f,  -->
<!--                     f = f, alpha = 1)|> -->
<!--   group_by(from_id, Care_Category) |> summarize(V_ic = sum(V_ij, na.rm = TRUE))  -->

<!-- test3 <- ttm_care_walk_15 |> filter(Care_Category == "Elder_Centric") |> -->
<!--   sp_avail_detailed(o_id = from_id, -->
<!--                     d_id = to_id,  -->
<!--                     pop = Population,  -->
<!--                     opp = f, -->
<!--                     r = f,  -->
<!--                     f = f, alpha = 1)|> -->
<!--   group_by(from_id, Care_Category) |> summarize(V_ic = sum(V_ij, na.rm = TRUE))  -->

<!--  test4 <- ttm_care_walk_15 |> filter(Care_Category == "Errand_Centric") |> -->
<!--   sp_avail_detailed(o_id = from_id, -->
<!--                     d_id = to_id,  -->
<!--                     pop = Population,  -->
<!--                     opp = f, -->
<!--                     r = f,  -->
<!--                     f = f, alpha = 1)|> -->
<!--   group_by(from_id, Care_Category) |> summarize(V_ic = sum(V_ij, na.rm = TRUE))  -->

<!-- test5 <- ttm_care_walk_15 |> filter(Care_Category == "Health_Centric") |> -->
<!--   sp_avail_detailed(o_id = from_id, -->
<!--                     d_id = to_id,  -->
<!--                     pop = Population,  -->
<!--                     opp = f, -->
<!--                     r = f,  -->
<!--                     f = f, alpha = 1)|> -->
<!--   group_by(from_id, Care_Category) |> summarize(V_ic = sum(V_ij, na.rm = TRUE))  -->

<!-- test <- rbind(test1, test2, test3, test4, test5) -->

<!-- access_15 <- access_15 |> left_join(test, by=c("from_id", "Care_Category")) |> ungroup() -->

<!-- rm(test, test1, test2, test3, test4, test5) -->
<!-- ``` -->