### 2024-09-03 WGiT/WGiT-pres.Rmd
Draft WGiT Presentation and edits to manuscript

---
title: "Towards caring 15-minute neighbourhoods"
subtitle: "Proximity Planning Congress 2024"
author: "Anastasia Soukhov, Léa Ravensbergen, Lucía Mejía Dorantes, Antonio Páez"
date: "09-10-2024"
output:
  xaringan::moon_reader:
    seal: false
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current% / 10" 
      highlightStyle: github
      highlightLines: true
      ratio: '16:9'
      countIncrementalSlides: false
      beforeInit: "macros.js"
---
```{r xaringan-themer, include=FALSE, warning=FALSE}
#this themer package allows us to change the .css themes through this chunk. Easier than changing the .css directly.
library(xaringanthemer)
style_mono_accent(
  base_color = "#7A003C", #maroon
  link_color = "#5E6A71", #grey
  text_bold_color = "#FDBF57", #gold
  text_slide_number_color = "#FDBF57", #gold
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```

```{r load-packages, include=FALSE, cache=FALSE}
options(htmltools.dir.version = FALSE)
library(RefManageR)

library(dplyr)
library(sf)
library(ggplot2)
library(tidyverse)
library(units)
library(patchwork)
library(ggspatial)
library(ggforce)
# library(geomtextpath)
library(knitr)
library(kableExtra)
library(here)
#install.packages("emo")
library(emo)
options(scipen = 999)
```

```{r bib-setup, include=FALSE, cache=FALSE}
BibOptions(check.entries = FALSE, bib.style = "authoryear", style = "markdown",
           dashed = TRUE, max.names = 1)
bib <- ReadBib(paste0(here(), "/bibliography/bibliography.bib"))
```

```{r sp_avail-function,include=FALSE}
#defining the spatial availability function, this prints all the intermediates
sp_avail_detailed <- function(x, o_id, d_id, pop, opp, r, f, alpha = 1){

  o_id <- rlang::enquo(o_id)
  d_id <- rlang::enquo(d_id)
  pop <- rlang::enquo(pop)
  opp <- rlang::enquo(opp)
  r <- rlang::enquo(r)
  f <- rlang::enquo(f)

  sum_pop <- x %>%
    dplyr::distinct(!!o_id,
                    .keep_all = TRUE) %>%
    dplyr::mutate(sum_pop = !!r*(!!pop)^alpha) %>%
    dplyr::pull(sum_pop) %>%
    sum()

  f_p <- dplyr::pull(x, !!r) * dplyr::pull(x, !!pop)^alpha / sum_pop

  sum_impedance <- x %>%
    dplyr::group_by(!!d_id) %>%
    dplyr::summarize(sum_impedance = sum(!!f))

  x <- x %>%
    dplyr::left_join(sum_impedance,
                     by = rlang::as_name(d_id))

  f_c <- dplyr::pull(x, !!f) / x$sum_impedance

  x$f_c <- f_c
  x$f_p <- f_p

  sum_pa <- x %>%
    dplyr::group_by(!!d_id) %>%
    dplyr::summarize(sum_pa= sum(f_p * f_c))

  x <- x %>%
    dplyr::left_join(sum_pa,
                     by = rlang::as_name(d_id))
  f_t <- (f_p * f_c) / dplyr::pull(x, sum_pa)

  x %>%
    dplyr::mutate(V_ij = !!opp * !!f_t)
}
```

class: center

# Towards caring 15-minute neighbourhoods

```{r, echo=FALSE, out.width='30%', fig.align='center'}
knitr::include_graphics(here("figures/mobility-of-care.png"))
```

#### Anastasia Soukhov (McMaster University)
#### Léa Ravensbergen (McMaster University), Lucía Mejía Dorantes (Karlsruhe, Germany), Antonio Páez (McMaster University)

#### _7th TRB International Conference on Women and Gender in Transportation: 10 September, 2024_

---

### The 15-Minute City

- a normative conceptualisation gaining ground in urban planning: reaching daily necessities/activities within 15\* minutes by walk, bike\*, and transit\*. `r Citep(bib, c("Allam2022","moreno_introducing_2021"))`

- Practical challenges:
  - Non-specific: 15-minutes? By which travel modes?
  - **For whom? To what destination?**

```{r, echo=FALSE, out.width='50%', fig.align='center'}
knitr::include_graphics(here("figures/15-m-city.jpg"))
```
_Shutterstock; Getty | Aerial views of Glasgow, Paris and Melbourne. Retrieved from Financial Times article "Welcome to the 15 Minute City" by N. Whittle in 2020_

???

Let's apply a gendered lens. The research objective: answering **For whom? To what destination?** with Mobility of Care

---

### The 15-Minute City and Mobility of Care

- Mobility of Care is _also_ a normative conceptualisation; coined in `r Citep(bib, c("sanchezdemadariagaMobilityCareIntroducing2013"))`, emphasizes all travel needed to sustain yourself and the household, i.e., all non-paid work related travel.

- Mobility of Care trips are often: under 15 minutes, proximate to home, and more frequently conducted by non-car modes (than work trips). Pair well with the 15-Minute City conceptualisation

- **Guiding research aim**: a method to identify what neighbourhoods are, are somewhat, and are not 15-minute _caring_ neighbourhoods

```{r, echo=FALSE, out.width='50%', fig.align='center'}
knitr::include_graphics(here("figures/graphs-trips-by-purpose.webp"))
```
_image from `r Citep(bib, c("sanchezdemadariagaMeasuringMobilitiesofCare2019"))`_

???

Mobility of Care... concerns everyone But as it stands, concerns women most.

- This type of travel has been traditionally understudied: 1) travel from home-to-work is often the focus of transportation planning, 2) travel surveys to inform this planning often obscure and other mobility of care travel.

---
### Case study: Hamilton, Canada

Research question: what neighbourhoods are, are somewhat, and are not 15-minute _caring_ neighbourhoods?

```{r, echo=FALSE, out.width='50%', fig.align='center'}
knitr::include_graphics(here("figures/pop_descriptive_plot_onlypop.png"))
```

---
### Case study: Hamilton, Canada

Research question: what neighbourhoods are, are somewhat, and are not 15-minute _caring_ neighbourhoods?

```{r, echo=FALSE, out.width='50%', fig.align='center'}
knitr::include_graphics(here("figures/care_categories_plot.png"))
```

---

### Quantifying 15-minute caring neighbourhoods

.pull-left[

#### A) Spatial accessibility of care destinations

How _many_ care destinations can be reached within a 15 minute walk? The _potential of opportunities for interaction_ `r Citep(bib, "hansenHowAccessibilityShapes1959")`.

  - Dependent-centric care destinations
  - Errand-centric care destinations
  - Grocery-centric care destinations
  - Health-centric care destinations

$S_i = \sum_j O_j f(c_{ij})$

]

.pull-right[

#### B) The diversity of categories

What is the _mix_ of care destination types within a 15 minute walk? The entropy measure:

$D_{i} = \frac{-\sum_{c} S_i^{c} / \sum_{c}S^{c}_{i} \times ln(S_i^{c} / \sum_{c}S^{c}_{i} )}{ln(n_c)}$

]

- $i$ is a set of parcel point origin locations; $j$ is a set of destination locations.
- $O^_j$ is a number of opportunities at $j$.
- $c_{ij}$ is the travel cost between $i$ and $j$.
- $f(\cdot)$ is an impedance function of $c_{ij}$; within the cumulative opportunity approach, it is a binary function that takes the value of 1 if $c_{ij}$ is less than a selected value

???

---

### Quantifying 15-minute caring neighbourhoods

.pull-left[

#### A) Spatial accessibility of care destinations

```{r, echo=FALSE, out.width='90%',fig.align='center'}
knitr::include_graphics(glue::glue(here::here(), "/figures/percats_copp_plot.png"))
```

]

.pull-right[

#### B) The diversity of categories

```{r, echo=FALSE, out.width='90%',fig.align='center'}
knitr::include_graphics(glue::glue(here::here(), "/figures/all_cats_entropy_copp_plot.png"))
```

]

???

### Typology and Profiles of 15-minute caring neighbourhoods: clustering via Self Organizing Maps

- To combine the Spatial Accessibility measures and the Diversity measure, and to profile the neighbourhoods, SOM is implemented.

- SOM is a pattern recognition method which reduces the dimensionality of the data into clusters without relying on any assumptions about the data structure.

The SOM procedure:
1 ) five superclusters of the four $S_i^c$ scores and the diversity score $D_i$ are identified. The number of superclusters strike a balance between variance explained and supercluster interpretation in what we define as ranging from "not caring" to "very caring" and "not complete" to "complete".

2) The superclusters are profiled using a variety of socio-economic and demographic indicators that the mobility of care literature finds significant.

---

### Typologies of 15-minute caring neighbourhoods

```{r, echo=FALSE, out.width='55%',fig.align='center'}
knitr::include_graphics(glue::glue(here::here(), "/figures/SClabel_copp_plot.png"))
```

???

very caring (mean 4 to 18.6; total p50 46) - very complete (div=0.9; p50 0.90) (Cluster 2)
caring (mean 2.1 to 9; total p50 18) - complete (div=0.81; p50 0.82) (Cluster 1)
somewhat caring (0.2 to 6; total p50 13) - complete (div=0.71; p50 0.74) (Cluster 4)
somewhat caring (0.2 to 6; total p50 7) - not complete (div=0.36; p50 0.40) (Cluster 3)
not caring (0 to 4.7; total p50 3) - not complete (div=0.02; p50 0) (Cluster 5)

---

### Profiles of 15-minute caring neighbourhoods typologies

```{r, echo=FALSE, out.width='60%',fig.align='center'}
knitr::include_graphics(glue::glue(here::here(), "/figures/access_sc_profiles_renamed.png"))
```

- Lower income households tend to reside in $\color{Green}{\text{very completely caring}}$ or $\color{GreenYellow}{\text{completely caring}}$ neighbourhoods; 
- Those with a `r emo::ji("up_arrow")` proportion of children tend to reside in $\color{orange}{\text{somewhat caring - not complete}}$ neighbourhoods;
- Higher income households in DAs with a `r emo::ji("down_arrow")` of single-parent households tend to reside in $\color{red}{\text{not caring - not complete}}$ neighbourhoods.

???

All variables that are tested:  Income_bottom + Age_0to14 + One_parent_families + Vis_minority + Not_in_labforce+ Income_gini+ walk_towork+ Owner_hld_CHN+ Hld_ownership+ Tenant_subsi_hld+ Tenant_hld_CHH+ LICO

Of a list of socio-economic variables, ones that characterise the 5 typologies the most are: 1) % in bottom half of the household income distribution, 2) children population (0 to 14) %, 3) single-parent household %, and 4) Low Income Cutoff (LICO) %

Who has resides in $\color{Green}{\text{very completely caring}$ or $\color{GreenYellow}{\text{completely caring}$ neighbourhoods?  (maybe people who cannot afford cars???)
- DAs that have lower median household incomes and a smaller children population (lower than 24% of the population are children 14 and under).
- and DAs which have especially high LICO are often in $\color{Green}{\text{very completely caring}$ neighbourhoods.

Who more often resides in $\color{BurntOrange}{\text{somewhat caring - not complete}$ neighbourhoods?  
- Of bottom income DAs, DAs with a high proportion of children population (higher than 24% of the population are children 14 and under).

Who more often resides in $\color{BrickRed}{\text{not caring - not complete}$ neighbourhoods?  
- DAs of higher income DAs and with lower proportion of single-parent households (13% or lower)

---

### Conclusions: classifying and profiling caring 15-minute neighbourhoods

.pull-left[

#### What we did:

- 15 minute-city is bridged with "Mobility of Care": 15-minute caring neighbourhoods.
- A data-driven pattern recognition method (Self-organizing maps) to create **15-minute caring neighbourhood** typologies using **accessibility** and **diversity** indicators is presented.
- Using Self-Organizing Maps again; each parcel tagged with a neighbourhood typology was profiled based on all relevant census variables.
]

.pull-right[

#### What we found and future considerations:

- **Gentrification concerns**: populations residing in completely caring neighbourhoods are often lower-income and vice-versa.

- **Concerning caring for children and children's self-care**: Lower-income DAs with a `r emo::ji("up_arrow")` proportion of children, and DAs with `r emo::ji("up_arrow")` proportion single-parent households have medium complete care access, why?

]

???