### 2024-10-10 editorial correspondance/Highlights.qmd
Draft to submit to journal. And some repo clean-up.

---
title: "Highlights"
format: html
---

Highlights:

- The 15-Minute City have emerged in popular urban planning and public discourse, emphasizing urban access to amenities by foot. However, critiques of the concept highlight how it falls short in addressing the pre-existing structural forces that drive inequalities.

-	The Mobility of Care framework is introduced as a way to clarify this critique of the 15-Minute City from the perspective of amenity focus. It argues that conventional, masculinist planning focuses on capital-driven amenities (work and leisure), neglecting essential "Care Destinations" like trips to drop-off children at daycare, grocery shopping, errands, attending medical appointment. These are pertinent to all people: but especially for women, and women who are lower-income.

- Through the empirical examination of Hamilton, Canada, accessibility to care destinations and the diversity of this accessibility is calculated for all residential parcels. Machine learning techniques are then used to classify areas of the city by their extent of _caring 15-minute neighbourhoods_ potential and create residential profiles along socio-economic dimensions.