### 2024-10-10 paper/Manuscript.qmd

---
title: "Caring 15-minute neighbourhoods"
author:
  - name: Anastasia Soukhov 
    email: soukhoa@mcmaster.ca
    affiliations:
        - name: McMaster University
          department: School of Earth, Environment & Society
          city: Hamilton, Canada
    attributes:
        corresponding: true
  - name: Léa Ravensbergen
    email: ravensbl@mcmaster.ca
    affiliations:
        - name: McMaster University
          department: School of Earth, Environment & Society
          city: Hamilton, Canada
  - name: Lucía Mejía Dorantes
    email: lucia.mejia.dorantes@gmail.com
    affiliations:
        - name: Consultant
          city: Karlsruhe, Germany
  - name: Antonio Páez
    email: paezha@mcmaster.ca
    affiliations:
        - name: McMaster University
          department: School of Earth, Environment & Society
          city: Hamilton, Canada
format: 
  # docx
  elsevier-pdf:
    keep-tex: true
    include-in-header:
      text: |
        \usepackage{graphicx}
        \usepackage{unicode-math}
        \usepackage{hyperref}
        \def\tightlist{}
        \usepackage{setspace}
        \doublespacing
        \usepackage{lineno}
        \linenumbers
    journal:
      #name: "Transportation Research Part A: Policy and Practice"
      formatting: preprint
      model: 3p
      cite-style: authoryear
keywords: ["Accessibility", "Mobility of Care", "Gender", "15-minute neighbourhood", "Care Complete Neighbourhoods"]
bibliography: bibliography.bib
---

```{r}
#| label: setup-knitr-options
#| include: false

knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      cache=FALSE)

```
```{r }
#| label: set-up-copy-bibliography
#| include: false

# WARNING: DO NOT MODIFY THE BIBLIOGRAPHY ON THIS FOLDER, WORK WITH THE ONE IN FOLDER ./bibliography/
file.copy(glue::glue(here::here(), "/bibliography/bibliography.bib"), 
          getwd(),
          overwrite = TRUE)
```

```{r}
#| label: function for creating comments

word_comment <- function(comment, highlight = "") {
  if (isTRUE(knitr:::pandoc_to() == "docx")) {
    paste0('[', comment, ']{.comment-start id="0" author="Anastasia"',
           'date="2024-05-16T15:00:00Z"}', highlight, '[]{.comment-end id="0"}')
  }
}
```

```{r library-setup}
library(biscale)
library(cowplot)
library(magick)
library(dplyr)
library(flextable)
library(ftExtra)
library(glue)
library(here)
library(mapview)
library(pander)
library(scales)
library(sf)
library(tidyr)
library(tmap)
library(tmaptools)
library(gt)
#install.packages("gtExtras")
library(gtExtras)
library(tidyr)
```

```{r}
load(glue::glue(here::here(), "/data/care_dest.rda"))
load(glue::glue(here::here(), "/data/access_15_wide.RData"))
load(glue::glue(here::here(), "/data/df_DA.RData"))
```

```{r reformat care-dest specific categories}
care_dest <- care_dest |> 
  mutate(Care_Category_Specific = case_when(
    Care_Category_Specific  == "Daycare_EarlyON" ~ "Daycare or EarlyON",
    Care_Category_Specific == "Community_Centre" ~ "Community or Recreation Centre",
    Care_Category_Specific == "Daycare_Centre" ~ "Daycare or EarlyON",
    Care_Category_Specific == "Recreation_Centre" ~ "Community or Recreation Centre",
    Care_Category_Specific == "School" ~ "School",
    Care_Category_Specific == "Long-Term_Care_Home" ~ "Long-Term Care or Retirement Home",
    Care_Category_Specific == "Retirement_Home" ~ "Long-Term Care or Retirement Home",
    Care_Category_Specific == "Senior_Centre" ~ "Senior Centre",
    Care_Category_Specific == "Seniors_Active_Living_Centre" ~ "Senior Centre",
    Care_Category_Specific == "Post_Office" ~ "Post Office",
    Care_Category_Specific == "Library" ~ "Library",
    Care_Category_Specific == "Banks" ~ "Bank or ATM",
    Care_Category_Specific == "Grocer_ConvenienceStore" ~ "Convenience Store",
    Care_Category_Specific == "Clinics" ~ "Hospital or Clinic",
    Care_Category_Specific == "Grocer_Retail" ~ "Grocer",
    Care_Category_Specific == "Dentist" ~ "Dentist",
    Care_Category_Specific == "Hosptial" ~ "Hospital or Clinic",
    Care_Category_Specific == "Hospital" ~ "Hospital or Clinic",
    Care_Category_Specific == "Pharmacy" ~ "Pharmacy",
    Care_Category_Specific == "Park" ~ "Park"),
  Care_Category = case_when(
    Care_Category == "Child_Centric" ~ "Dependent-centric",
    Care_Category == "Elder_Centric" ~ "Dependent-centric",
    Care_Category == "Errand_Centric" ~ "Errand-centric",
    Care_Category == "Grocery_Centric" ~ "Grocery-centric",
    Care_Category == "Health_Centric" ~ "Health-centric"))

care_dest <- care_dest |> rename("care_ID" = "ID") |> mutate(care_ID = as.character(care_ID)) 
# care_dest|> st_drop_geometry() |> group_by(Care_Category) |> summarise(count = n())
```

```{r creating boxplot, eval=FALSE}
summary_boxplots <- access_15_wide |> select(c(scluster, starts_with("norm_"), div_allcats)) |>
  mutate(scluster = case_when(scluster == 1 ~ "A+",
                              scluster == 2 ~ "A",
                              scluster == 3 ~ "A-",
                              scluster == 4 ~ "B-",
                              scluster == 5 ~ "B",
                              scluster == 6 ~ "B+",
                              scluster == 7 ~ "D"),
         scluster = factor(scluster, ordered = TRUE,
                           levels = c("A+", "A", "A-","B+","B","B-","D")))

# summary_boxplots_Aplus <- summary_boxplots |> filter(scluster == "A+") |> select(-c("scluster"))
# summary_boxplots_A <- summary_boxplots |> filter(scluster == "A") |> select(-c("scluster"))
# summary_boxplots_Aminus <- summary_boxplots |> filter(scluster == "A-") |> select(-c("scluster"))
# summary_boxplots_Bplus <- summary_boxplots |> filter(scluster == "B+") |> select(-c("scluster"))
# summary_boxplots_B <- summary_boxplots |> filter(scluster == "B") |> select(-c("scluster"))
# summary_boxplots_Bminus <- summary_boxplots |> filter(scluster == "B-") |> select(-c("scluster"))
# summary_boxplots_D <- summary_boxplots |> filter(scluster == "D") |> select(-c("scluster"))

summary_boxplots_tidy <- summary_boxplots |>
  select(c("scluster","div_allcats",
           "norm_tot_copp_isc_norm_dep", "norm_tot_copp_isc_norm_err",
           "norm_tot_copp_isc_norm_groc", "norm_tot_copp_isc_norm_health")) |>
  rename("1"="div_allcats",
         "4"="norm_tot_copp_isc_norm_dep", 
         "2"="norm_tot_copp_isc_norm_err",
         "5"="norm_tot_copp_isc_norm_groc", 
         "3"="norm_tot_copp_isc_norm_health") |>
  pivot_longer(!scluster, names_to="variables", values_to = "normalized")
```

```{r, eval=FALSE}
box_plot <- ggplot(summary_boxplots_tidy, 
       aes(x=variables, y=normalized, fill=scluster)) +
  geom_boxplot(outlier.shape = NA,
               linewidth = 0.1) +
  scale_fill_manual(values=c("A+" = "#1A9850",
                               "A" = "#91CF60", 
                               "A-" = "#D9EF8B", 
                               "B+" = "#FEE08B",
                               "B" = "#FDAE61",
                               "B-" = "#F46D43",
                               "D" = "#D73027"),
                    name = " ") +
    scale_x_discrete(labels = c("1" = "Category Totals",
                                "2" = "ERRAND TOTALS",
                                "3" = "HEALTH TOTALS",
                                "4" = "DEPENDENT TOTALS",
                                "5" = "GROCERY TOTALS")) +
  coord_flip()+
  theme_bw() +
  theme(axis.title.y=element_blank()) +
  ylab("Min-max normalized values") 
box_plot
ggsave(file= file.path(here("figures/box-plot.png")))
```

```{r creating-table2-pt1}
summary_df_ALL <- access_15_wide |> select(c(starts_with("total_"), div_allcats)) |>
  group_by() |>
  summarise(across(everything(), list(mean = mean, sd = sd))) %>%
 mutate_if(is.numeric, round, digits =1) |>
  mutate_all(as.character)
summary_df_ALL <- cbind(data.frame(scluster = "TOTAL"),summary_df_ALL)

summary_df <- access_15_wide |> select(c(scluster, starts_with("total_"), div_allcats)) |>
  mutate(scluster = case_when(scluster == 1 ~ "A+",
                              scluster == 2 ~ "A",
                              scluster == 3 ~ "A-",
                              scluster == 4 ~ "B-",
                              scluster == 5 ~ "B",
                              scluster == 6 ~ "B+",
                              scluster == 7 ~ "D")) |>
  group_by(scluster) |>
  summarise(across(everything(), list(mean = mean, sd = sd))) %>%
 mutate_if(is.numeric, round, digits =1) |>
  mutate_all(as.character)

merge_every_two_columns <- function(df, num_rows) {
  # Loop through every second column starting from column 2
  for (col in seq(2, ncol(df), by = 2)) {
    if (col < ncol(df)) {  # Ensure the column has a next one to merge with
      for (i in 1:num_rows) {
        df[i, col] <- paste0(df[i, col], " (", df[i, col + 1], ")")
      }
    }
  }
  return(df)
}

summary_df <- merge_every_two_columns(summary_df, 7) |>  select(-contains("sd"))
summary_df_ALL <- merge_every_two_columns(summary_df_ALL, 1)  |>  select(-contains("sd"))
summary_df <- rbind(summary_df, summary_df_ALL)

summary_df <- summary_df |>
  pivot_longer(!scluster) |>
  pivot_wider(names_from =scluster )

summary_df_ALL <- summary_df_ALL |>
  pivot_longer(!scluster) |>
  pivot_wider(names_from =scluster )

summary_df$name <- c("GROCERY TOTALS", "Convenience Store", "Grocer",
                     "DEP. TOTALS", "Comm. or Rec. centre", "Daycare or EarlyON","LTC or retirment home","Park","School","Senior centre",
                     "HEALTH TOTALS","Dentist","Hospital or clinic","Pharmacy",
                     "ERRAND TOTALS","Bank or ATM","Library", "Post office",
                     "Category diversity")

summary_df <- summary_df[,c(1,3,2,4,6,5,7,8,9)]
```

```{r, echo=FALSE}
df_DA_nogeom <- df_DA |> st_drop_geometry()
df_DA_nogeom <- df_DA_nogeom[,30:81]

df_DA_nogeom <- df_DA_nogeom |> mutate(
  Income = `v_CA21_907: Median after-tax income of household in 2020 ($)`,
  Avg_numb_of_kids = `v_CA21_498: Average number of children in census families with children`,
  Not_in_labforce = `v_CA21_6504: Not in the labour force`/(`v_CA21_6498: Employed` + `v_CA21_6501: Unemployed` + `v_CA21_6504: Not in the labour force`),
            Unemployed = (`v_CA21_6501: Unemployed`)/(`v_CA21_6498: Employed` + `v_CA21_6501: Unemployed` + `v_CA21_6504: Not in the labour force`),
            Income_gini = `v_CA21_1142: Gini index on adjusted household after-tax income` ,
            Vis_minority = `v_CA21_4875: Total visible minority population`/ (`v_CA21_4914: Not a visible minority`+ `v_CA21_4875: Total visible minority population`),
            walk_towork =  `v_CA21_7647: Walked`/ (`v_CA21_7647: Walked` + `v_CA21_7635: Car, truck or van` + `v_CA21_7644: Public transit` + `v_CA21_7650: Bicycle` + `v_CA21_7653: Other method` ),
            bike_towork =  `v_CA21_7650: Bicycle` / (`v_CA21_7647: Walked` + `v_CA21_7635: Car, truck or van` + `v_CA21_7644: Public transit` + `v_CA21_7650: Bicycle` + `v_CA21_7653: Other method` ),
            pt_towork = `v_CA21_7644: Public transit`/ (`v_CA21_7647: Walked` + `v_CA21_7635: Car, truck or van` + `v_CA21_7644: Public transit` + `v_CA21_7650: Bicycle` + `v_CA21_7653: Other method` ),
            car_towork =  `v_CA21_7635: Car, truck or van`/ (`v_CA21_7647: Walked` + `v_CA21_7635: Car, truck or van` + `v_CA21_7644: Public transit` + `v_CA21_7650: Bicycle` + `v_CA21_7653: Other method` ),
            Owner_hld_CHN = `v_CA21_4308: % in core housing need (57)` ,
            Hld_ownership = `v_CA21_4306: % of owner households with a mortgage (58)` ,
            Hld_sheltercosts = `v_CA21_4307: % of owner households spending 30% or more of its income on shelter costs (55)` ,
            Tenant_sheltercosts = `v_CA21_4315: % of tenant households spending 30% or more of its income on shelter costs (55)` ,
            Tenant_subsi_hld = `v_CA21_4314: % of tenant households in subsidized housing (61)` ,
            Tenant_hld_CHH =`v_CA21_4316: % in core housing need (57)` ,
            Low_or_no_deg = (`v_CA21_5868: No certificate, diploma or degree` + `v_CA21_5871: High (secondary) school diploma or equivalency certificate`) / (`v_CA21_5868: No certificate, diploma or degree` + `v_CA21_5871: High (secondary) school diploma or equivalency certificate` + `v_CA21_5874: Postsecondary certificate, diploma or degree`))
```
```{r, echo=FALSE}
cors_df_DA_nogeom <- cor(df_DA_nogeom, use = "complete.obs") |> as.data.frame()
#cors_df_DA_nogeom[cors_df_DA_nogeom <0.7 | cors_df_DA_nogeom == 1] <- ""

One_parent_family_inspect <- cors_df_DA_nogeom |> select(One_parent_families)
LICO_inspect <- cors_df_DA_nogeom |> select(LICO)
Income_inspect <- cors_df_DA_nogeom |> select(Income)
```

```{r population in each bar column for the Decision Tree, eval=FALSE}
df_DA |> filter(`v_CA21_907: Median after-tax income of household in 2020 ($)` <= 68750) |>  select("Population") |> st_drop_geometry()  |> sum(na.rm=T) / df_DA |>  select("Population") |> st_drop_geometry()  |> sum(na.rm=T)

df_DA |> filter(`v_CA21_907: Median after-tax income of household in 2020 ($)` > 68750 & `v_CA21_907: Median after-tax income of household in 2020 ($)` <  91500) |>  select("Population") |> st_drop_geometry()  |> sum(na.rm=T) / df_DA |>  select("Population") |> st_drop_geometry()  |> sum(na.rm=T)

df_DA |> filter(`v_CA21_907: Median after-tax income of household in 2020 ($)` >=  91500) |>  select("Population") |> st_drop_geometry()  |> sum(na.rm=T) / df_DA |>  select("Population") |> st_drop_geometry()  |> sum(na.rm=T)
```

# Abstract

The "15-Minute City" concept has been embraced by global leaders to promote human-scale neighbourhoods with transport and land-use designs that support short trips to daily necessities. This paper bridges the 15-Minute City to "Mobility of Care",  a framework that foregrounds travel to care destinations, travel done predominately by women. This focus contrasts the more commonly studied travel to employment and leisure destinations. While the 15-Minute City concept is flexible enough to consider all destination types, gendered examinations are relatively lacking in the literature, and little research to date has focused explicitly on care destinations. This gap is addressed in this paper by identifying which areas in the city of Hamilton, Canada are ‘caring 15-minute neighbourhoods’. To do so, a database of care destinations was compiled to estimate the number (using the cumulative opportunities accessibility measure) and diversity of mix (using the entropy measure) of care destinations within a 15-minute walk from residential parcels. Using data-driven machine learning techniques (Self-Organizing Maps and Decision Trees), neighborhoods were classified into 'caring 15-minute neighbourhood' typologies that are examined across residential socio-economic profiles. Our results suggest that the majority of caring 15-minute neighbourhoods are in the urban core, where more lower-income households reside. In contrast, not caring 15-minute neighbourhoods in higher-income peripheral areas. Areas that make good candidates for urban policy intervention are identified, and the implications of enhancing 15-minute walkable caring access are discussed in relation to gender mainstreaming in transportation planning.

Keywords: Accessibility; 15-Minute Neighborhoods; Chrono-Urbanism; Cities of Proximity; Mobility of Care; Inequality; Gender Gap;

# Introduction

The "15-Minute City" has been recently adopted by leaders as a way to promote human-scaled cities [@teixeiraClassifying15minuteCities2024]. As introduced in @moreno_introducing_2021, the 15-Minute City is a urban planning model based in chrono-urbanism, a theory that emphasizes the positive impact on quality of urban life when urban space becomes multi-rythmic, in opposition to the segregation of urban time and space for individual uses and mobility [@mulicekUrbanRhythmsChronotopic2015; @morenoVilleQuartHeure2016]. It is in opposition to the forces that supported the formation of single-use neighbourhoods such as industrial Fordism and single-use zoning [@mulicekUrbanRhythmsChronotopic2015; @morenoVilleQuartHeure2016]. The "15-Minute City" refers to a city where relevant destinations within a walkable (or bikeable or transit supportive) radius are reachable by all. This form of neighbourhood planning would allow individuals to reclaim time spent on car mobility, giving way to sustainable modes and prompting urban spaces that are responsive to human needs and environmental sensibilities [@Allam2022].

Though the term 15-Minute City is relatively new, neighbourhood planning practice aiming to foster community and local travel to amenities is not. As recent examples, planned neighbourhoods dominated post-WWII built urban form: the Neighbourhood Unit Concept in the western world [@brodyNeighbourhoodUnitConcept2013] and a parallel concept of the “mikrorayons” neighbourhoods in the Soviet Union  [@kissfazekasCircleParadigms15minute2022]. Along different dimensions, these planning forms have been extensively critiqued [@talenSocialSciencePlanned2017]. In this respect, neighbourhood planning approaches such as the 15-Minute City concept are being re-embraced by some, but from bottom-up perspective using co-creation tools and meaningful resident participation to achieve equitable and justice neighbourhood forms [@mahmoudCocreationPathwayUrban2021]. How to equitably prescribe urban form that 'enables' contact with social opportunities (instead of 'engendering it' from the top-down) is still under debate. Questions remain, including: what destinations really matter?, what mode and travel time threshold is most appropriate?, and who benefits? For instance, these dimensions are under discussion among proximity-based planners who use accessibility-based tools [@silvaProximitycentredAccessibilityConceptual2023; @silvaRegionalAccessibilityUndermining2022; @guzmanProximityEnoughCritical2024].

From another perspective, feminist geographers, urban planners and other researchers agree that urban structure influences travel, but city planning falls short in planning for a neutral identity. In contrast, cities ought to be planned according to the multiple identities of the inhabitants [@vacchelli_towards_2018; @urbandevelopmentviennaGenderMainstreamingUrban2013]. Researchers have demonstrated how mobility patterns differ according to gender [@lawWomenTransportNew1999; @cresswell_gendered_2008; @levy2013travel; @little_gender_1994; @tronto_toward_1990]. Gendered social norms and cultural factors play a role in how, where, with whom, what for, and how far we move. Women have higher activities related to care [@ilo_care_2018], no matter at what life stage they are [@garciaromanGenderDifferencesTime2022]. For all these reasons the spatial organization of activities has a different impact according to gender, the 15-Minute City concept should not be gender blind.

This work builds a theoretical bridge between the 15-minute city and mobility of care concepts to answer questions about what destinations matter and who benefits from proximity planning. The Mobility of Care term was coined in @sanchezdemadariagaMobilityCareIntroducing2013; it refers to all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands.  While decades of research have examined types of household-serving trips (e.g., shopping trips, escorting to school), @sanchezdemadariagaMobilityCareIntroducing2013 was the first to consider them all as one category. In doing so, Mobility of Care is shown to be a significant proportion of daily travel, characterizing the trip purpose of approximately 30% of adults’ trips [@sanchezdemadariagaMobilityCareIntroducing2013; @sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory]. We argue Mobility of Care is especially relevant to the 15-Minute City, as care trips are necessary for _all_ people and they most often occur at the local level. Moreover, care trips take on tremendous gendered and socio-economic importance, as women, and especially those from lower income households, complete the majority of care trips [@ravensbergen2023exploratory]. Despite care travel's importance in advancing gender equality -and equity more broadly-, mobility of care and access to care destinations have been under-examined relative to employment destinations. This is especially pertinent in accessibility research that has largely focused on employment points of interest e.g., [@farberOntarioLineSocioeconomic2019; @duarteInfluenceJobAccessibility2023; @ryanAccessibilitySpaceTime2023].

The objectives of this research are two-fold. First, we aim to theoretically bridge Mobility of Care with the 15-Minute City concept to re-imagine what local amenities matter and to define an associated accessibility and diversity of accessibility measures. Second, we take a data-driven approach to classify areas into 'caring 15-minute neighbourhoods' typologies and examine associated residential profiles in an empirical examination of the city of Hamilton, Canada.

# Review of neighbourhood planning literature

## From the 15-Minute City to the NUC

In the last decades, the need for more sustainable, healthier, and livable cities have received significant attention. Planners and decision-makers have proposed alternatives to mobility-based neighbourhood planning principles to ones that center proximity to urban functions [@pozoukidou15MinuteCityDecomposing2021]. In this context, the 15-Minute City is now in the public spotlight [@logan_xminute_2022; @moreno_introducing_2021]. The 15-Minute City was introduced in @moreno_introducing_2021 along with four dimensions: density (in terms of people per km$^2$), diversity (including mixed-land use and diversity of people), the temporal and spatial proximity to essential services, and digitalisation. However, the framework presented has been criticized within academic and planning circles [@guzmanProximityEnoughCritical2024]. Critiques highlight how the concept falls short in addressing the pre-existing structural forces and individual characteristics that drive inequalities, influencing who benefits or could benefit the most from such a paradigm [@dimarino15minuteCityConcept2023; @willberg15minuteCityAll2023]. A 15-Minute City for all people is an aspirational ideal, but how will existing mobility and accessibility inequalities be addressed. Without directed and context-specific solutions, this popular concept risks becoming an empty city branding exercise. [@pozoukidou15MinuteCityDecomposing2021; @gowerPlanningInnovationCity2022].

Reflecting the flexible and aspirational presentation of the 15-Minute City; cities that have adopted the concept have done so using a diverse range of definitions and tools along with indistinct universal approaches. A trail blazer was Portland, US in the Portland Plan [@city_of_portland_20minute_2010] of April 2012. Adopted before the '15-Minute City' concept of today, it aims to foster an inclusive city development based on prosperity, education, health, and equity over a 30 year horizon. The plan aims to promote neighbourhood self-sufficiency and connectivity to city centers and centres of employment. The progress report describes a high-level focus on equitable service delivery to all residents with equity concerning topics related to racial equity and people with disabilities [@portland_government_portland_2017], similarly taking on an "all populations" approach. Later, other cities adopted proximity-based goals using similar neutral approaches. For example, the 15-Minute City plan announced during the Paris mayor Anne Hidalgo’s 2020 re-election campaign, which helped popularize the concept in today's public discourse, emphasized six key social functions that should be easily accessible from any location [@ville_de_paris_paris_2022]. These locations included: housing, work, health care, groceries, education and leisure. The 15-Minute City concept inspired language in the agendas of other cities in the Western world such as Ottawa, in the Canadian context, who also adopted a 15-Minutes approach in their recent Official Plan [@ville_dottawa_quartier_2021]. @teixeiraClassifying15minuteCities2024's global review finds that the 15-Minute City concept is in early phases of implementation around the world and the diverse range of definitions, strategies, and instruments present a significant knowledge and implementation gap. In other words: the 15-Minute City is aspirational but how do we get there and who will benefit?

The past can shed light on cautionary tales. While the 15-Minute City concept is new, neighbourhood planning to improve society outcomes is not: in fact the literature has drawn parallels from the 15-Minute City to Clarence A. Perry's Neighbourhood Unit Concept (NUC) from the 1930s [@kissfazekasCircleParadigms15minute2022]. The NUC is a socio-spatial normative scheme widely adopted by government agencies (and the real-estate community) in the Western world after the Second World War [@talenSocialSciencePlanned2017; @solow1969concept]. Pairing well with the objectives of planning agencies at the time, Perry's NUC would allow for efficient mass-building of cellular units that prioritized the perceived functional needs of women and children: each unit providing proximate access to an elementary school and supporting community facilities, shopping, parks and housing [@talenSocialSciencePlanned2017; @brodyNeighbourhoodUnitConcept2013]. The NUC primarily prioritized local service provision, though Perry had confidence in good design's contribution to 'neighbourhood spirit' [@hall2014cities]. From the NUC's conception until the end of the 1960s, planners' aspirational attempts to prescribe social meaning to the neighbourhood's physical form had been criticized to near extinction [@talenSocialSciencePlanned2017]. By social scientists, the overestimation on the impact of built environment on social life was critiqued. By other planners, the inability to agree on the specificity of the neighbourhood (i.e., population size and the type, quality, and quantity of amenity) as well as how neighbourhood units are connected to other neighbourhoods and the rest of the region.  In response to these criticisms, neighbourhood planning proponents redefined deterministic terminology; their prescriptive physical form will 'enable' social contact with opportunities rather than 'engendering' it.

The redefined bottom-up approach to community and neighbourhood planning has been adopted by New Urbanists, from which the 15-Minute City stems [@kissfazekasCircleParadigms15minute2022]; however, the effectiveness of these bottom-up processes remain to be seen due to the contemporary nature of the concept. Though, related research can be examined. In recent years, the question of _how can physical form be planned to enable an improved quality of life, for whom, and with what outcomes_ has occupied the urban planning research agenda. For instance, the examination of low-income households access to transportation and their likelihood of gaining employment [@blumenbergDriveWorkRelationship2017; @bastiaanssenDoesBetterJob2022] or the relationship between children's access to public transit and participation in after-school activities [@palmRolePublicTransit2020]. A new wave of researchers and practitioners focused on local and context-specific relationships with the proximity to destinations have also emerged [@silvaProximitycentredAccessibilityConceptual2023; @silvaRegionalAccessibilityUndermining2022]. As reviewed in the city plans that have adopted 15-Minute City approaches, the NUC's criticisms, and recent urban planning research, the question of _how to enable improved quality of life through urban built environment_ has shown to be highly context-sensitive, with the need for further investigation.

## Tools: accessibility methods, diversity measures and typology-classification

In examining how to enable improved quality of life through urban built environment, accessibility measures have become an increasingly popular tool. These measures are a way to quantify the ease of reaching destinations from a given point in space and have been used to examine urban areas through just and sustainable city planning agendas [@vale2023accessibility; @handyAccessibilityIdeaWhose2020]. The 15-Minute City is an amenity-provision neighbourhood planning concept aimed at enabling the creation of urban environments that enhance life quality, making it well-suited for analysis using accessibility methodologies [@guzmanProximityEnoughCritical2024]. Recent works have applied accessibility measures to investigate the 15-Minute City across different geographic scopes. For instance, in Naples, Italy [@gaglioneUrbanAccessibility15minute2022], Barcelona, Spain [@graellsCityCitiesMeasuring2021], Vancouver, Canada [@hosford15minuteCityReach2022], and in urban areas across Europe [@vale2023accessibility]. The "cumulative opportunity" measure has been applied in many 15-Minute City examinations. This measure quantifies how many destinations can be reached from a point in space within a given travel time threshold, pairing well with examinations that have a normative travel time threshold, like the 15-Minute City. Furthermore, the cumulative opportunity measure is widely appreciated for its intuitive computation and popularity among practitioners [@handyAccessibilityIdeaWhose2020; @handyMeasuringAccessibilityExploration1997; @chengInvestigatingWalkingAccessibility2019]. However, accessibility measures other than the cumulative opportunity have also been applied, reflecting the diversity of measures in the literature [@guzmanProximityEnoughCritical2024].

Another method that compliments the assessment of the 15-Minute City regarding the diversity of opportunity types is entropy. The entropy measure, based on the Shannon-Wiener index of species diversity, expresses relative evenness within a sequence [@whittaker1972evolution]. In urban studies and planning literature, the flexible concept of entropy has been used to characterize land-use mix diversity [@frankLinkingObjectivelyMeasured2005; @ewingTravelBuiltEnvironment2010a], particularly in the context of non-work trips [@cerveroTravelDemand3Ds1997], walking as a mode [@luUrbanDensityDiversity2017; @mavoaIdentifyingAppropriateLanduse2018], and suburban sprawl [@randallGISBasedLand2015] as relevant to this work. It has also been used to understand mobility behaviour [@mcbride2020exploration; @montero2023applying; @montero2023role]. However, work incorporating entropy indices alongside accessibility analysis are not exceptionally common. There are examples of their use as parameters within accessibility scores for employment opportunities [@chengMeasuringUrbanJob2013; @daiIncorporatingJobDiversity2018] and, more recently, to describe the diversity in transit facilities [@yinIncorporatingFacilityDiversity2024]. Within the conceptualisation of the 15-Minute City, diversity of land-use is a key feature. In this way, the entropy measure theoretically complements the 15-Minute City.

Classification algorithms also show promise as a tool in the assessment of the 15-Minute City, as they have been useful in identifying mobility and spatial typologies within the transportation planning literature. Often, these algorithms take the form of machine learning approaches: for instance, @burkeGeospatialAnalysisFramework2022 uses spatially constrained multivariate clustering, to develop urban form typologies related to the 15-Minute City. The k-means algorithm is another popular method: @ganUnderstandingUrbanMobility2020 applies it to analyse travel behaviour and classify metro stations based on mobility patterns. Similarly, [@montero2023role; @montero2023applying] cluster travel behavior profile typologies representing different population segments and place of residence in the Barcelona Metropolitan Area. Combing classification techniques is also common: for instance, @delmelleDifferentiatingPathwaysNeighborhood2017 uses self-organizing maps (SOM) algorithm to group U.S. neighbourhoods by minimizing dissimilarity in attributes before applying k-means for further classification. Similarly, @victorianoTimeSpaceMoney2020 uses SOM to classify individuals' travel patterns according to the dissimilarity in mobility attributes and uses the Decision Tree algorithm to partition the data into interpretable classifications.

## Mobility of Care and feminist 15-minute neighbourhoods

Rather than focusing on _all_ destinations, it may valuable to examine those related to caring activities. This re-framing offers a feminist perspective on which urban functions matter, connecting well with the 15-Minute City concept. Caring activities, which fulfill the physical, psychological, and emotional needs of others, are among the most essential and fundamental activities in society [@ilo_care_2018]. Yet, they are the most unequal, undervalued, and even devalued activities worldwide. Conventionally, caring activities have been borne on women’s shoulders [@haydenGrandDomesticRevolution1982; @hochschild_second_2012]. According to @ilo_care_2018, women and girls perform more than three-quarters of the total amount of unpaid care worldwide, a gender gap that varies geographically [@RN5]. This unequal share of caring responsibilities leads to multifaceted gendered differences: in career development, profession selection, contract type, pay gap, and time poverty, as recognized by various international organisations [@EIGE2016; @ilo_care_2018]. In terms of spatial and transportation planning, almost one third of daily trips are for care purposes [@sanchezdemadariagaMobilityCareIntroducing2013; @sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory]. From this research motivation, Sánchez de Madariaga coined the term Mobility of Care in 2013 to refer to all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands [@sanchezdemadariagaMobilityCareIntroducing2013]. While an undercurrent of research had examined these unique household-serving trips over the decades, her seminal work was the first to consider all these trips as one category and demonstrate how mobility of care is a significant proportion of daily travel.

The Mobility of Care concept also explicitly integrates inter-sectional equity considerations, a common criticism leveled at the 15-Minute City [@guzmanProximityEnoughCritical2024]. Perhaps unsurprisingly given the gendered division of care work discussed, women have consistently been found to complete more mobility of care trips than men. In one study, mobility of care comprised 32% of women’s daily trips compared to 28% of men’s. While this gap is significant, it was found to be far greater in lower income households where women complete 20% more care trips than men [@ravensbergen2023exploratory]. Sánchez de Madariaga not only shows how important these mobility of care trips are, but also highlights the ways in which “mobility of care is systematically under-represented in any analysis of urban transport” (p. 37). As a product of the masculinst bias in planning, transport surveys and tools often do not directly capture mobility of care, which re-enforces the idea that these trips are not a significant part of daily mobility. In this respect, the feminist perspective of the cities of proximity is still underestimated with only few examples on the topic [@gil_sola_choose_2022; @macintyre_her_2022].

Pairing well with the focus on shorter trips and the potential use of sustainable modes within the 15-Minute City, Mobility of Care trips are more local, shorter-distance and trip-chained. Compared to the trip to work, mobility of care trips are more frequently completed by foot, and less frequently by transit or bicycle [@ravensbergen2023exploratory]. However, little work to date examines walking to care destinations through the Mobility of Care framework. Though there is ample literature that examines walking to care destinations, such as schools (e.g., [@RN7;@RN8;@RN9]) and grocery stores (e.g., [@RN10;@RN11]), they tend to consider singular care destinations in research focusing on walkability.

In summary, we reviewed the neighborhood planning framework from which the 15-Minute City concept stems and is situated within. We then reviewed tools common and useful in investigating the 15-Minute City. Finally, we introduced the Mobility of Care concept, theoretically linking it to the 15-Minute City to highlight the importance of care destinations for all but especially along gendered and socio-economic lines. The objectives of this work are two-fold, achieved through an empirical examination of Hamilton, Canada, a mid-sized city:

- First, by theoretically bridging Mobility of Care with the 15-Minute City concept, we re-imagine what local amenities matter and calculate associated accessibility and entropy measure for care destinations.
- Second, using a bottom-up data-driven approach, we apply machine learning methodologies to classification areas into _15-minute caring neighbourhoods_ and investigation their associated residential profiles along socio-economic lines.

<!--  the objectives of the paper are completed as follows: -->

<!-- 1)	The cumulative accessibility measure and the entropy measure are used to quantify both the level of access and the diversity of access to care destinations within a 15-minute walk. -->
<!-- 2) Based on these accessibility and diversity outputs, "Completely Caring 15-Minute Neighbourhood" typologies are generated using the SOM algorithm.   -->
<!-- 3)	A Decision Tree is then applied to these typologies and socio-economic variables of populations who reside in these neighbourhoods. This is done to delineate data-driven profiles on which groups have the potential to benefit from what Completely Caring 15-Minute Neighbourhood.  -->

# Data

## Case study context

This work focuses on Hamilton, Ontario, a mid-sized city on the shore of Lake Ontario. Hamilton has a heterogenous land-use, with a populated and dense urban core, surrounded by suburban development, which is itself surrounded by rural communities. The Niagara Escarpment runs through Hamilton, and results in a city with two key elevations: a more dense lower city that contains the downtown core and the elevated suburban development referred to as ‘the Mountain’. In this work, we analyse the residential parcel centroids, 143,882 locations. We aggregated the points at the level of Canadian Census Dissemination Area (DA) along with the population and population per parcel plots in @fig-Fig1.

```{r plot-figure1, out.width=600, out.height=600}
#| label: fig-Fig1
#| fig-cap: "The number of residential parcels per DA in 2020 (top), the population (middle) retrieved from the 2021 Canadian Census, and the rate of population per parcel per DA (bottom). All scales in quartiles. Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/figures/pop_descriptive_plots_1.png"))
```

Hamilton also exhibits spatial disparities in social and economic indicators; their spatial distribution is visualised in @fig-Fig2. The densely populated inner city is characterised by lower average incomes, and a higher prevalence of households living under the low-income cutoff thresholds (LICO). Notably, the suburban areas of the city tend to have a greater proportion of children and a lower proportion of one parent households.

```{r plot-figure2, out.width=600, out.height=600}
#| label: fig-Fig2
#| fig-cap: "Socio-economic and demographic variables that characterise accessibility to care destinations retrieved from the 2021 Canadian Census. All scales in quartiles. Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/figures/pop_descriptive_plots_2.png"))
```

\newpage

## Care destination dataset

A spatial dataset of care destinations for Hamilton was compiled as detailed in ( _work removed for double-blind review_ ). The dataset includes 14 types of care destinations that were classified into four categories: dependent-centric (e.g., the destinations for child- and elder-centric escorting trips), grocery-centric, health-centric, and errand-centric. Notably, these categories were generated following the travel purpose categories created in the mobility of care research by @sanchezdemadariagaMeasuringMobilitiesCare2019. Care categories, sources of data, and preparation notes are detailed in @tbl-Tbl1. The spatial distribution of destination type are visualised in @fig-Fig3 by their category.

```{r creating-table-content}
dat <- data.frame(matrix(c("Depedent-centric",
                           "Schools, daycares, and community centres, recreation centres, parks, senior centres, long-term care homes, and retirement homes: 1,265 locations are included.",
                           "[@opendatahamiltonEducationalInstitutions2022; @governmentofontarioOntarioDataCatalogue2023; @cityofhamiltonChildCareRegistry2023; @opendatahamiltonParksDatabase2022; @opendatahamiltonRecreationCommunityCentres2022; @ontarioministryofhealthgeohubMinistryHealthService2023]",
                           "Grocery-centric",
                           "Convenience stores and grocery stores (e.g., large retailers as well as speciality food grocers, health food grocers): 381 destinations are included.",
                           "[@axledataConsumerData2023]",
                           "Health-centric",
                           "Hospitals, pharmacies, clinics, and dentist offices: 421 destinations are identified.",
                           "[@ontarioministryofhealthgeohubMinistryHealthService2023; @hamiltonniagarahaldimandbranthealthlineWalkInMedicalClinics2023]",
                           "Errand-centric",
                           "Libraries, post offices, and banks: 158 destinations are identified.",
                           "[@opendatahamiltonLibraries2022; @axledataConsumerData2023; @canadapostPostOfficeLocator2023; @bankofmontrealBankMontrealBranch2023; @thehongkongandshanghaibankingcorporationlimitedhsbcHSBCBranchABM2023; @nationalbankNationalBankBranches2023; @royalbankofcanadaRBCBranchABM2023; @scotiabankScotiabankABMBranch2023; @thetorontodominionbankTDBankBranch2023]."
                           ), ncol = 3, byrow = T))
```

```{r creating-table, ft.arraystretch = 1}
#| tbl-cap: "Description of care destinations categories, notes on data preparation and associated data sources."
#| label: tbl-Tbl1
set_flextable_defaults(font.size= 8,
                       font.family = NULL,
                       text.align = 'justify',
                       table.layout = 'fixed')

flextable(dat) %>%
  ftExtra::colformat_md() %>%
  compose(
    i = 1, j = c(1,2,3), part = "header", value = as_paragraph(c("Care category", "Care destination types", "Sources"))) %>%
  width(j = 1, width = 1.7, unit = "cm") %>%
  width(j = 2, width = 10, unit = "cm") %>%
  width(j = 3, width = 5.3, unit = "cm")|>
  flextable::line_spacing(space = 0,
                          part = "body")
```

```{r plot-figure3, out.width=600, out.height=600}
#| label: fig-Fig3
#| fig-cap: "The locations of care destinations in Hamilton separated by the author-generated categories of: dependent-, errand-, grocery- and health- centric care categories. Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/figures/care_categories_plot.png"))
```

\newpage

## Travel time to care destinations estimations

Travel behaviour to care-oriented destinations is often uncounted and hence comprehensive travel times to all destination types in @tbl-Tbl1 is unavailable. To overcome this gap, travel time from residential parcel locations in Hamilton and care destinations in @fig-Fig3 are approximated assuming travel by foot at an average speed (3.6 km/hr). This estimation is done using the 'travel_time_matrix()' function from the {r5r} package [@pereiraR5rRapidRealistic2021] using R version 4.3.2 [@coreR2023]. The inputs into the function are: the locations of 143,882 residential parcels as origins, the 2,225 care locations as destinations, and a OpenStreetMap road network including walking infrastructure [@geofabrikOntarioCanadaOpen2023]. In line with the 15-Minute City, a maximum walking travel time of 15 minutes is specified and an origin-destination travel time matrix of the shortest travel time from origin to destination is calculated. The resulting matrix contains 2,014,502 rows, representing walking travel times from each parcel to reachable care destinations within 15 minutes.

# Methods

The following sub-sections detail the methods to classify Hamilton into spatial degrees of _caring 15-minute neighbourhood_. First, accessibility to each of the 14 destination types from each of the 143,882 residential parcel locations is described. Second, the entropy measures used to calculate the diversity of accessibility to each of the care categories is detailed. Third, we describe how accessibility and diversity values for each parcel are input into a Self-Organizing Map (SOM) algorithm, and how the resulting clusters are analyzed using a Decision Tree algorithm to narrate residential profiles based on socio-economic variables. In sum, this methodology presents a data-driven approach to examine what neighbourhoods in a city have the potential to provide 15-minute caring access, at what level of intensity and completeness, and the socio-economic characteristics of those who are most benefited or burdened.

## Accessibility: the cumulative opportunity measure

To capture the quantity of access to each type of destinations, a cumulative opportunity accessibility score $S^t_i$ is calculated. Scores for each of the 14 care destination types $t$ is calculated for every parcel $i$. The calculation takes the following mathematical form:

$$
S^t_i=\sum_{j}O_j^{t}\cdot f(c_{ij})
$$
\noindent Where:

- $i$ is a set of parcel point origin locations.
- $j$ is a set of care destination locations of type $t$.
- $O^t_j$ is a number of opportunities of category type $t$ at $j$.
- $c_{ij}$ is the travel cost between $i$ and $j$.
- $f(\cdot)$ is an impedance function of $c_{ij}$; within the cumulative opportunity approach, it is a binary function that takes the value of 1 if $c_{ij}$ is less than a selected value [@handyMeasuringAccessibilityExploration1997].
- $S_{i}^t$ is the cumulative opportunity accessibility score, the sum of weighted opportunities reachable within $f(\cdot)$, at each $i$ for each $t$.

## Diversity in opportunity accessibility: the entropy measure

The entropy measure, as defined in @cerveroTravelDemand3Ds1997, is used to represent the diversity of care destination accessibility. For each parcel, a value between 0 to 1 is calculated, where 1 represents total evenness in the number of care opportunities in each category that can be reached. The mathematical formulation takes the following form:

$$
D_{i} = \frac{-\sum_{t} [S_i^{t} / \sum_{t}S^{t}_{i} \times ln(S_i^{t} / \sum_{t}S^{t}_{i} )]}{ln(n_t)}
$$

\noindent Where:

- $i$ is a set of parcel point origin locations.
- $t$ is a set of care destination types (e.g., school, grocery, park, etc.)
- $n_t$ is the count of care destination types $t$. In this work, this value is 14.
- $S_{i}^t$ is the cumulative opportunity accessibility score, the sum of weighted opportunities reachable within a 15-minute walk from $i$.
- $D_i$ is the diversity score.

Notably, $D_i$ represents evenness in the type of care categories a parcel can access. For example, if a parcel has an access score of $S_i^t = 0.5$ for all destination types, it will receive a diversity score of $D_i = 1$, just as if it had an access score of $S_i^t = 10$ for each destination. Conversely, a parcel may be assigned a low $D_i$ score if its accessibility scores differ across categories, regardless of whether those scores are low or high.

## Machine learning classification: SOM and Decision Trees

We use two machine learning techniques in this work. First, SOM is an unsupervised technique implemented to reduce the data dimensionality and create interpretable clusters related to the intensity and completeness of caring access. This is done by imputing each parcel as an observation with its associated accessibility and diversity attributes, and the data being rearranged onto a two-dimensional space based on its minimizing dissimilarity in its neighbourhood. An appropriate number of superclusters are selected and assigned labels based on the quantity and diversity of care access provided, i.e., the degree by each the parcel is located in a _caring 15-minute neighbourhood_. Then, a Decision Tree is run to characterise the socio-economic profiles of who resides in neighbourhoods associated with the superclusters. Together, this combined approach leverages the unsupervised data-driven classification power of the SOM with the interpretation of Decision Trees. The procedure used in this work is similar to that used in @victorianoTimeSpaceMoney2020. In this work, rather than each observation representing an individual's daily mobility behaviour (with associated variables), each observation represents a parcel location with calculated care accessibility and accessibility diversity scores.

For the SOM step, the function 'trainSOM()' from {SOMbrero} R package is used [@VillaVialaneix2017]. The input variables include the 143,882 parcels, each as individual observation along with 15 variables: the 14 calculated accessibility scores $S_i^t$, normalized to the min-max range score within each $t$, and one diversity value $D_i$. Otherwise, defaults for all other parameters are assumed, relying on the data-driven heuristics embedded in the 'trainSOM()' function. Consequently, a 100 node (10 by 10 grid) SOM structure using euclidean distance and square topology is produced. Simply put, the SOM algorithm proceeds as follows: a 2D grid of nodes is created as specified by the analyst, where a node represents a point in the reduced-dimensional space. Upon initialization, each node is assigned a random weight vector of the same dimension as the input data (in our case, 15). From the input data, a random observation with its associated weight vector (i.e., one parcel point with 15 variables) is selected and the Euclidean distance between its weight vector and all nodes in the grid are calculated. The node with the smallest distance (i.e., the smallest dissimilarity) is labelled the 'best matching unit' as it is the node that best represents the input observation. After this best matching unit is identified, its own weight and its neighbouring nodes are updated to become more like the input observation. The process of finding best matching units and updating their weights is repeated for every observation, multiple times, until the results converge. As mentioned, this competitive learning process produces a 100 node SOM structure where each observation (parcel) is assigned to 1 node with an associated dissimilarity index. The SOM output is typically examined through a dissimilarity dendogram and an associated dissimilarity variance explained plot to select an appropriately representative 'superclusters' [@VillaVialaneix2017; @victorianoTimeSpaceMoney2020].

For the Decision Tree step, the supercluster-classified parcels identified in the SOM step are used as _labels_ and socio-economic and demographic indicators related to the Mobility of Care literature are used as _features_. Features are retrieved from the 2021 Canadian Census [@governmentofcanadaCensusPopulation2023]. This step creates residential profiles of the superclusters, allowing us to explore the characteristics of residents within the superclusters in a data-driven way. To estimate the Decision Tree, the 'rpart()' function in the {rpart} R package is used [@Therneau2023]; default parameters for classification splitting along with each value being weighted by the population present in the associated DA is assumed. The Decision Tree algorithm is a supervised learning technique that begins by splitting a subset of the input data into branches based on a selected feature with the lowest impurity score (i.e., the least mixing of label membership within a branch). This process is recursively repeated for each data subset, selecting the next best feature. Ultimately, the data is classified into distinct classes, with class membership explained by traversing the branches defined by the features that characterise the partitions within the Decision Tree. Notably, the absence of features from the Decision Tree does not necessarily imply they are irrelevant for classification, rather, they are less relevant than other features. Put another way, when considering features that are highly correlated, such as income level and LICO, not all relevant variables may be present in the tree [@victorianoTimeSpaceMoney2020].

# Results

## Overview of access to care destinations
How many care destinations can be reached within a 15-minute walk is summed for each of the four care categories defined in @tbl-Tbl1. The median parcel value for each DA is visualised in @fig-Fig4 and the median parcel diversity of care destinations accessible is presented in @fig-Fig5. Together, @fig-Fig4 and @fig-Fig5 represent summaries of 15 variables that served as inputs into the SOM algorithm.

```{r plot-figure4, out.width=600, out.height=600}
#| label: fig-Fig4
#| fig-cap: "The number of care destinations that can be reached within a 15-min walk per care category for a median parcel in each DA. The values are a summary of the 14 acessibility scores that served as inputs into the SOM. Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/figures/percats_copp_plot.png"))
```

```{r plot-figure5, out.width=450, out.height=450}
#| label: fig-Fig5
#| fig-cap: "The diversity measure based on the proportion of care category accessibility in Figure 4. These values are also a summary of the 15th input into the SOM. Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/figures/all_cats_entropy_copp_plot.png"))
```

\newpage

## Identification of completely caring 15-minute neighbourhood typologies

In @fig-Fig4, the concentration of opportunities for all category types is centered in the downtown of Hamilton, with the highest concentration being in the lower city (near the lake shore) and pockets of the high concentration further from the shoreline. Grocery-centric destinations appear to be the most concentrated, followed by health-centric and dependent-centric caring destinations. Errand-centric destinations are the most sprawled. In many ways, @fig-Fig4 mirrors the spatial distribution of care destinations (@fig-Fig3) as the range of 15-minutes of travel is small. In @fig-Fig5, areas that have high care accessibility tend to have high diversity as well, though there are exceptions in pockets of the city outside the downtown core that have higher diversity but low levels of accessibility for all care categories. Similarly, there are areas with low diversity within the downtown core that have only moderate or high accessibility to certain care destination types.

Based on the SOM methodology discussed, 7 superclusters are identified from the 10 by 10 grid of SOM nodes. As diagnostics for the selection of the number of superclusters, the dissimilarity-index-based dendrogram with the proportion of parcels represented in each supercluster alongside the variance explained plot is visualised in @fig-Fig6. Labels representing grades A to D qualifying the 7 superclusters are then assigned. These grades as also reflected in @fig-Fig6 and represent the quantity of accessibility and diversity in accessibility per care category. Higher grades (A+ and A) corresponding to the highest accessibility and diversity, while lowest grades (D) representing lowest accessibility and diversity scores.

```{r plot-figure6, out.width = '100%', out.height = '100%'}
#| label: fig-Fig6
#| fig-cap: "The 7 resulting superclusters from the SOM the 10 by 10 grid output represented in a dendrogram (left) and the proportion of variance unexplained by cluster (right)."
knitr::include_graphics(glue::glue(here::here(), "/figures/all_access_sc_plot_small.png"))
```

\newpage

The grade for each supercluster was carefully assigned by the authors by referring to descriptive statistics and their notable trends. These statistics are summarised in @tbl-Tbl2 and in the boxplots in @fig-Fig7.

- A+ and A have exceptionally high caring accessibility for all destination types and high diversity scores in the top quantile or above. Together, these grades represent 24% of all parcels in the city.
- A- has high caring accessibility, but low diversity. Interestingly, A- is like A but with much higher dependent-centric destinations, particularly parks, schools, and daycares, with only moderately high scores for all other destination types. This disbalance lowers its overall diversity score. However, it can be characterized as a supercluster that demonstrates potential in being retrofitted to provide A+ or A level of completely caring access. 4% of all parcels are represented by the A- grade. 
- B+ and B superclusters present about average completely caring access. These superclusters can serve as the benchmark for what 'average' completely caring access in Hamilton currently looks like. These grades represent 15% of all parcels.
- B- provides above average dependent-centric destination access, particularly parks, daycares and schools, but below average access to other destination types and hence has low diversity scores. B- is like A- as it demonstrates complete caring 15-minute potential if retrofitted. As these parcels demonstrate caring access to some destinations, they may have the potential to be retrofitted to support complete access to all caring destination types. This grade represents 19% of all parcels.
- D superclusters demonstrate the lowest scores all-around, representing room for land-use improvement that addresses complete and caring 15-minute access. This supercluster characterizes the largest number of parcels, representing 38% of parcels in the city.

```{r creating-table2-pt2, ft.arraystretch = 1}
#| tbl-cap: "Mean and (standard deviation) of each SOM classified cluster by input variable and additional summary variables. Variables included in the SOM algorthim are in regular case, while additional summary variables are indicated by ALL CAPITAL LETERS."
#| label: tbl-Tbl2
#|
# summary_df %>%
#   gt() %>%
#   tab_header(
#     title = " ",
#   ) %>%
#   cols_label(
#     name = " "
#   ) %>%
#   gt_highlight_cols(
#     2:3, fill = "green", alpha = 0.3
#     ) %>%
#   gt_highlight_cols(
#     4, fill = "green", alpha = 0.2
#     ) %>%
#   gt_highlight_cols(
#     5:6, fill = "orange", alpha = 0.3
#     ) %>%
#     gt_highlight_cols(
#     7, fill = "orange", alpha = 0.2
#     ) %>%
#   gt_highlight_cols(
#     8, fill = "red", alpha = 0.3
#     ) %>%
#   tab_style(
#     style = cell_borders(
#       sides = c("top"),
#       color = "black",
#       weight = px(1.5),
#       style = "solid"
#     ),
#     locations = cells_body(rows = c(1,4,11,15,19))
#   ) |>
#   tab_options(
#     table.font.size = 10,
#     table.layout = 'fixed',
#     table.align = 'center',
#     table.width = pct(100)
#  )

set_flextable_defaults(font.size= 8,
                       font.family = NULL,
                       text.align = 'justify',
                       table.layout = 'fixed')

summary_df <- summary_df |> rename(" " = "name")

flextable(summary_df) %>%
 ftExtra::colformat_md() %>%
  flextable::width(j = 1, width = 3.06, unit = "cm") %>%
  flextable::width(j = 2:9, width = 1.53, unit = "cm") %>%
  bg(j= 2, bg = "#48B376") |>
  bg(j= 3, bg = "#91CF60") |>
  bg(j= 4, bg = "#D9EF8B") |>
  bg(j= 5, bg = "#FFFFBF") |>
  bg(j= 6, bg = "#FEE08B") |>
  bg(j= 7, bg = "#FDAE61") |> 
  bg(j= 8, bg = "#D75C56") |>
  hline(i = c(3,10,14,18))  |>
  flextable::line_spacing(space = 0,
                          part = "body")

# RColorBrewer::brewer.pal(7, "RdYlGn")
# "#D73027" "#FC8D59" "#FEE08B" "#FFFFBF" "#D9EF8B" "#91CF60" "#1A9850"
```

```{r plot-figure7, out.width=800, out.height=300}
#| label: fig-Fig7
#| fig-cap: "A boxplot demonstrating summary variables and category diversity that define the superclusters."
knitr::include_graphics(glue::glue(here::here(), "/figures/box-plot.png"))
```

\newpage

To spatially demonstrate where the superclusters are located, representative supercluster labels for each DA are visualised in @fig-Fig8. This visualisation is created by grouping parcels by their DA, calculating the median supercluster label for each label, and selecting the label that is most dominant within that DA. For reference, the median number of parcels in each DA is `r df_DA$parcel_count |> median() |> round()` and supercluster label membership within a DA is typically pure, with the median membership being `r df_DA$max_label_val |> median() |> scales::percent()` of a single cluster.

```{r plot-figure8, out.width=450, out.height=450}
#| label: fig-Fig8
#| fig-cap: "The maximum median parcel supercluster membership per DA. Escarpment is visualised as a grey dashed line. Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/figures/SClabel_plot.png"))
```

The resulting superclusters visualised in @fig-Fig8 are a combination of the cumulative opportunity and diversity of access measures translated into meaningful typologies through the SOM methodology. To reiterate, accessibility is the measure of _potential_ interaction, e.g., how many care destinations one could reach within a 15-minute walk, these typologies are useful in identifying which areas of the city are providing regionally high levels of complete caring access, which areas are regionally average, and which areas are regionally below average. Furthermore, neighbourhoods which provide disbalanced access can be further investigated.

In @fig-Fig8, it can be observed that the excellent and exceptional neighbourhoods (A, A+), which are very caring and very complete 15-minute neighbourhoods, are located within the center of the city and closest to the shoreline in the downtown core of Hamilton. The above and below average neighbourhoods (B+ B) are often proximate to the excellent and exceptional neighbourhoods and are more prevalent in south of the escarpment within the center Hamilton. The escarpment is a physical barrier, with few pedestrian-accessible access points to traverse; hence the typologies describing neighbouring DAs separated by the escarpment are often different. Below average (D) neighbourhoods are located in peripheral areas outside the center of Hamilton, in areas where urban form is characterised by lower density residential housing and auto-dependent mobility.

Furthermore in @fig-Fig8, A- and B- grades stand out as offering high access for children-centric destinations for their grade-group but below grade-group average access in other destination types. These neighbourhoods may be more suitable for populations who prioritize walkable access to destinations like schools, parks and daycares, and find access to other types of caring destinations less important. These neighbourhoods also stand out as demonstrating potential to be retrofitted to provide more complete access if additional destination types were located within their neighbourhoods.

## Profiles of who does and does not reside in caring 15-minute neighbourhoods

To enhance the meaning of the superclusters beyond a descriptive and spatial conceptualisation of "caring" and "complete",  _who?_ resides in what neighbourhood is investigated through the Decision Tree results. The input features of the Decision Tree are the supercluster labels and the feature variables are various population-weighted socio-demographic characteristics of the 2021 Canadian Census, namely: median household income, % below the median household income, % LICO prevalence, average number of children per household, % population aged 0 to 14, % not in the labour force, % not employed, Gini index on adjusted household after-tax income, % visible minority, % single parent household, % who walk to work (relative to bike, care/truck/van, public transit and 'other'), % of owner household in core housing need (i.e., inadequate housing structure or paying higher than 30% income on housing), % of tenant households in subsidized housing, % of tenant households in core housing need, % no certification or with only a highschool diploma.

Of all the included input variables, _median household income_ proved to be the most meaningful in partitioning the superclusters data. @fig-Fig9 provides the Decision Tree with the significant splits in median household income and the proportion composition of supercluster along each branch for three terminal Decision Tree nodes. While the algorithm is unable to homogenously use each supercluster, @fig-Fig9 is helpful to report a narrative of who may reside in what caring/complete supercluster. Particularly, the Decision Tree demonstrates a more pure supercluster membership for only two of the three terminal Decision Tree nodes: supercluster A+ (exceptional completely caring access) and D (not caring and complete access). For reference, the city-wide mean household income is `r df_DA_nogeom$Income |> mean(na.rm=T)|> round() |> prettyNum(big.mark = ",") ` (SD: `r df_DA_nogeom$Income |> sd(na.rm=T)|> round() |> prettyNum(big.mark = ",")` and median: `r quantile(df_DA_nogeom$Income,0.5, na.rm=T) |> unname()  |> prettyNum(big.mark = ",")`). The notable splits for income are >\$91,500, >\$68,750, or between \$68,750 and \$91,500; roughly split by lower income, middle income, and higher income median household brackets.

<!-- - % single parent household (median: `r df_DA$One_parent_families |> median(na.rm=T) |> scales::percent()` (`r df_DA$One_parent_families |> sd(na.rm=T) |> scales::percent()`) ), and  -->
<!-- - % prevalence of LICO (median: `r df_DA$LICO |> median(na.rm=T) |> round() |> paste0("%")` (`r df_DA$LICO |> sd(na.rm=T) |> round() |> paste0("%")`) ) -->

```{r plot-figure9, out.width=600, out.height=450}
#| label: fig-Fig9
#| fig-cap: "The Decision Tree demonstrating median household income splits and superclusters composition within each branch."
knitr::include_graphics(glue::glue(here::here(), "/figures/access_sc_profiles_justincome.png"))
```

Describing @fig-Fig9, the bar column on the right represents DAs with higher household income. In these DAs, the majority of parcels are labeled as D superclusters, i.e not caring or complete 15-minute neighbourhoods. Within this higher-income-representing bar column, the second highest proportion of supercluster membership are B- parcels. B- parcels are low in most amenity types except a few child-centric destinations and show promise in being more easily improved than the D parcels. The higher income column represents 39% of the population, the largest proportion of any of the three columns. The left column in @fig-Fig9 corresponds to the lowest income, i.e., below 68,750 CAD median household income, and accounts for 32% the population. It is dominated by A+ parcels along with A and A- parcels, in the largest quantities relative to other bar columns, however, proportions of all other superclusters are also present. The middle column is defined by parcels with a mix of supercluster classifications and represents DAs with middle households incomes, between 68,750 CAD and 91,500 CAD representing 28% of the population.

Though median income was the most useful in partitioning the parcels by their supercluster labels in @fig-Fig9, other variables are still important. Namely, the following variables in order of importance and correlation coefficient with the median household income variable in brackets, are listed: the % below the median household income (`r Income_inspect[c("Income_bottom"),] |> round(digits=2)`), % single parent household (`r Income_inspect[c("One_parent_families"),] |> round(digits=2)`), % no certification or with only a highschool diploma (`r Income_inspect[c("Low_or_no_deg"),] |> round(digits=2)`), the average number of children per household (`r Income_inspect[c("Avg_numb_of_kids"),] |> round(digits=2)`), % not in the labour force (`r Income_inspect[c("Not_in_labforce"),] |> round(digits=2)`), % LICO prevalence (`r Income_inspect[c("LICO"),] |> round(digits=2)`), and % who walk to work (`r Income_inspect[c("walk_towork"),] |> round(digits=2)`). <!-- Of note, when median household income is removed from the Decision Tree, a similar structure is created, except with above average Single-Parent family and LICO prevalence splitting the data.-->

It is notable that median household income is highly or moderately correlated with many of these important variables, and discussing these correlations is useful for interpretation alongside the Decision Tree diagram. For instance, single-parent families are most positively correlated with the proportion of households being in the bottom income distribution (`r Income_inspect[c("Income_bottom"),] |> round(digits=2)`) and lower or no diploma (`r One_parent_family_inspect[c("Low_or_no"),] |> round(digits=2)`), and most negatively with the median after-tax household income (`r One_parent_family_inspect[c("v_CA21_907: Median after-tax income of household in 2020 ($)"),] |> round(digits=2)`), reflecting nationwide trends [@governmentofcanadaDailyMainHighlights2024; @governmentofcanadaPrevalenceLowIncome2024]. Though not all single-family households are below the LICO (i.e., single-family households are not highly correlated with LICO at the DA-level in Hamilton), DAs with higher concentrations of single-family households and LICO prevalence tend to have A+ access. Conversely, DAs with lower single-parent households are in DAs with a higher concentration of parcels with D access. These findings are notable from an equity perspective since economic disadvantaged often intersects with other socio-demographic characteristics.

# Discussion

Geographically, Hamilton is city that offers very caring - and completely caring - 15-Minute neighbourhoods for some, but not all. There are evident spatial inequalities, with areas ranging from excellent (A+ and A), average (B+ and B), to well below average (D). Some areas, like those labelled A- and B-, show potential for improvement within their grade groups. While the downtown core has the highest concentration of _caring 15-minute neighborhoods_, certain areas outside the core do as well. This finding somewhat contrasts employment accessibility studies, such as @elgeneidyNonstopEquityAssessing2016's assessment of public transit access to employment in the Greater Toronto Area (including Hamilton), where access is heavily concentrated in the downtown core, more so than appears to be present in this work. Though our paper uses a different methodology, both studies highlight distinct patterns: indeed, A+, A or A- care access is concentrated in the downtown core but it is also present in certain pockets outside of the center of Hamilton, leading to interesting conjectures. For instance, some of the A-grade neighborhoods in Hamilton likely follow land-use principles that emphasized residential proximity to amenities, similar to the NUC reviewed. The potential overlap of the NUC with caring 15-minute neighborhoods warrants further investigation.

Who currently resides in Hamilton's completely caring 15-Minute neighbhorhoods is also demonstrated to be a somewhat optimistic story. Parcels that provide A+ completely caring access tend to be in DAs that are economically disadvantaged. Economic disadvantage tends to intersect with other identities such as gender [@lightmanMeasuringEconomicExclusion2018]. And as reviewed in this work, all women and especially those from lower income households tend to complete most mobility of care trips [@ravensbergen2023exploratory]. Furthermore, lower-income households tend to also be single-parent households. Broadly, single-parent households are more likely to be time-disadvantaged [@nieuwenhuisSingleparentFamiliesWork2018], and tasked with a higher proportion of care duties [@craigTimeCareComparison2004]. In this way, the most economically disadvantaged groups having A+ complete and caring access is an optimistic finding. However, Hamilton is experiencing gentrification [@ellisyoungWeReJust2018]; rents are rising along the future light rail transit corridor and throughout the city [@vandermerweSpilloverGentrificationMidsized2021; @mayersLightTransitDocumenting2023]. Toronto, Hamilton's larger and higher-rent neighbouring city, is spilling gentrification into Hamilton's downtown core, (re)producing neighbourhoods based on Toronto's middle class identities [@mayersLightTransitDocumenting2023]. In these ways, the lower income residents of Hamilton's A+ neighbourhoods are more likely to be displaced, which is matter of wellbeing and justice. There are now few low-rent choices that provide the same exceptional level of access as the downtown core, hence lower household income residents that currently reside in A+ neighhourhoods will likely displaced in the coming years if current trends continue.

In discussing policy interventions that equitably increase completely caring 15-Minute neighbourhoods in Hamilton, this work's presents a methodology to create city-wide relative typologies and investigate who currently resides in what neighbourhood, as a stepping stone for further investigation. Neighbourhoods with the lowest grades (D) and with the highest potential in being improved (i.e., high accessibiltiy for certain types of destinations but not all) are neighbourhoods with B- and A- grades. However, the work's findings demonstrate a mismatch in household income and completely caring access: higher income areas are often the ones targeted for improvement. This raises important questions from the land-use policy perspective that need to be further analysed. For instance, is it equitable to focus policy on ameliorating neighbourhoods that are already higher-rent though they tend to be more rural, single-use zoned and car-dependent (parcels with D grades)? Further, of the parcels that provide high child-centrics but low otherwise, A- parcels (better access) tend to be in DAs with lower-income households more so than B- parcels (lower access than A-). From the perspective of ameliorating land-use to support the equitable distribution of 15-minute caring neighbourhoods which areas be targeted? Who should benefit, and how? If the policy initiative is targeted to specific neighbourhoods: sustainability linked to car dependency and equity are in tension. This harkens to what role a planned neighbourhood, and bottom-up planning approaches that include the evaluation of travel behaviour by socio-economic and demographic profile should be adopted to achieve equitable 15-minute cities.

As is the case for all research, the results should also be interpreted along with methodological assumptions. This work measures spatial accessibility which is a measure of potential interaction with all reachable destinations from an origin. These destinations, however, may not be relevant to people at an origin, e.g., they may be underutilized (walkable access to 2 schools may be underused, as children often only attend 1 school), the trip physically undesirable (the walk may be along an arterial with high traffic speeds, making the trip unlikely to ever happen by foot), or the average walking speed assumed may not reflect the walking speeds of all populations [@willberg15minuteCityAll2023]. Further, people who reside in these neighbourhoods may also disagree with the neighbourhood's label. The grade labels are region-relative (e.g., high accessibility in Hamilton may be subjectively insufficient for some) and they do not consider subjective perceptions that influence accessibility (e.g., though a neighbourhood has many opportunities, it may not feel safe to access them). Furthermore, accessibility is calculated from the point of residential parcels. Care trips are not necessarily completed from home, in-fact, they are often completed in complex trip-chains [@scheinerWomenComplexDaily2015]. In this way, the results flatten the dynamic patterns of care trips. These methodological assumptions should all be taken together when interpreting the results. In this way, the methodologyand findings presented identifies spatial and socio-economic trends that should be further investigated.

# Conclusion

This study makes three types of contributions to the transportation and city planning literature: empirical, methodological and theoretical. At the empirical level, areas of the mid-sized City of Hamilton have been typified by their degree of '15-Minute Caring Neighbourhood' potential. Methodologically, we applied the longstanding accessibility measure of cumulative opportunities and entropy to classify spatial areas based on how many destinations could be reached in a 15-minute walk and the diversity of destination type. These values were then clustered using a novel machine learning approach, SOM, to generate meaningful typologies for discussion and further comparison with socio-economic composition of the area. We find A+ and A 15-minute caring neighbourhoods are located within the downtown core and in certain suburban pockets of the city, while the peripheral regions provide D level caring access. A- and B- areas are also identified as neighbourhoods that already support a high amount of children-centric destination access, and could be improved to provide better complete care access. Theoretically, our work puts forth an explicitly caring 15-Minute Neighbourhod conceptualisation, bridging the Mobility of Care and the 15-Minute City concepts. We discuss how measuring caring neighbourhoods can be explicitly considered within city planning.

Our work is of relevance to researchers and practitioners planning equitable and sustainable cities. Instead of prescribing an urban form design principle, such as "all local amenities should be within a 15-minute walking distance", it instead examines an empirical example to determine which areas in the city have the _potential_ to be 15-minute neighbourhood based on the existing spatial accessibility offered by the urban environment and walking transport network. To this end, this data-driven methodology introduces a way to identify neighbourhoods that have potential, almost have potential, and are far from containing this potential to support future context-specific qualitative work.

# References

::: {#refs}
:::